---
title: "gganimate"
author: "Yifei Liu"
date: "3/17/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective: learn how to use gganimate package

```{r}
library(tidyverse)
detach("package:dplyr", unload=TRUE)
library(dplyr)
library(demography)
library(gganimate)
library(forecast)
library(scales)
library(foreign)
library(data.table)
library(transformr)
library(av)
theme_set(theme_minimal())
```


# gganimate live cookbook - Thomas Lin Pedersen

[gganimate live cookbook - Thomas Lin Pedersen](https://www.youtube.com/watch?v=3UY7Or0v_dI&list=WL&index=493)


transition_time

```{r}

anim <- p + transition_time(year)

anim

```

Label string interpolation


```{r}

p + transition_time(year) +
  labs(title = "Year: {frame_time}")

```

view_follow

```{r}
p + 
  transition_time(year) +
  labs(title = "Year: {frame_time}") +
  view_follow(fixed_y = TRUE)

```

shadow_wake 

```{r}

p + 
  transition_time(year) +
  labs(title = "Year: {frame_time}") +
  shadow_wake(wake_length = 0.1,
              alpha = FALSE)


```

shadow_mark

```{r}

p + 
  transition_time(year) +
  labs(title = "Year: {frame_time}") +
  shadow_mark(alpha = .3, size = .5)

```


temperature


```{r}
library(GSODR)

copenhagen <- nearest_stations(LAT = 55.676098,
                               LON = 12.568337,
                               distance = 50)

austin <- nearest_stations(LAT = 30.267153,
                           LON = -97.7430608,
                           distance = 50)

copenhagen_temp <- get_GSOD(2018, station = copenhagen)

austin_temp <- get_GSOD(2018, station = austin)

january <- bind_rows(copenhagen_temp, austin_temp) %>%
  dplyr::filter(MONTH == "1") %>%
  distinct(YDAY, STNID, .keep_all = TRUE) %>%
  mutate(location = if_else(
    STNID %in% copenhagen, "Copenhagen", "Austin"
  ))

select(january, YDAY, TEMP, location, STNID)



```
static version

```{r}

p <- ggplot(january) +
  geom_line(aes(YDAY, TEMP,
                color = location,
                group = STNID)) +
  scale_color_brewer(type = "qual") +
  theme(legend.position = "bottom") +
  labs(x = "Day of Month",
       y = "Temperature (celcius)",
       color = NULL)
p
```
transition_time


```{r}

p + 
  transition_time(YDAY)

```
transition_time + shadow_mark

```{r}

p + transition_time(YDAY) +
  shadow_mark()

```
transition_reveal

```{r}

p + transition_reveal(YDAY) +
  geom_point(aes(x = YDAY, y = TEMP,
                 color = location,
                 group = STNID))

```

```{r}

p + transition_reveal(YDAY) +
  geom_point(aes(x = YDAY, y = TEMP,
                 color = location,
                 group = seq_along(STNID)))

```



```{r}

p <- ggplot(january) +
  geom_smooth(aes(YDAY, TEMP,
                  color= location)) +
  scale_color_brewer(type = "qual") +
  theme(legend.position = "bottom") +
  labs(x = "Day of month",
       y = "Temperature (celcius)",
       color = NULL)
p
```

using stat()

```{r}

p + transition_reveal(stat(x))

```


# gganimate getting started - gganimate github example

[gganimate getting started](https://gganimate.com/articles/gganimate.html)


## get started

```{r}
library(gganimate)

p <- ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point()

p

```
my attempt

```{r}


q <- gapminder %>%
  mutate(decade = year %/% 10 * 10) %>%
  ggplot(aes(gdpPercap, lifeExp,
             size = pop, color = country)) +
  geom_point(alpha = .7, show.legend = F) +
  scale_color_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(x = "GDP per capita",
       y = "Life Expectancy")

q

```







```{r}

anim <- p + transition_states(Species,
                              transition_length = 2,
                              state_length = 1)

anim
```

my attempt

```{r}

anim_q <- q + 
  transition_states(decade,
                    transition_length = 4,
                    state_length = 1)

anim_q
```


transition_states() splits up plot data by a discrete variable and animates between teh different states

## Easing

```{r}

anim + 
  ease_aes('cubic-in-out')

```

```{r}

anim_q + 
  ease_aes('cubic-in-out')

```


ease_aes() defines the velocity with which aesthetics change during an animation

```{r}

anim +
  ease_aes("bounce-out")

```

## Labeling

```{r}

anim +
  labs(title = "Now showing {closest_state}",
       subtitle = "Frame {frame} of {nframes}")


```

my attempt

```{r}

anim_q +
  labs(title = "Now showing {closest_state}",
       subtitle = "Frame {frame} of {nframes}")

```

## Object permanence


```{r}

ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_line(aes(group = rep(1:50, 3)), color = "grey") +
  geom_point()

```


1. add some aesthetics taht distinguish the different speices

```{r}

ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point(aes(color = Species)) +
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

```


2. set the group directly


```{r}

ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point(aes(group = seq_along(Species)))+
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

```

the group aesthetic defines how the data in a layer is matches across the animation

In general 2. is prefered as it makes the intend explicit, it also make it possible to match data with different discrete aesthetics such as keeping our (now obviously faultly) transition while having different color for the different species


```{r}

ggplot(iris, aes(Petal.Width, Petal.Length))+
  geom_point(aes(color = Species, group = 1L)) +
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

```

my attempt

```{r}

gapminder %>%
  mutate(decade = year %/% 10 * 10) %>%
  ggplot(aes(gdpPercap, lifeExp,
             size = pop)) +
  geom_point(aes(color = continent, group = 1L),
             alpha = 0.7, show.legend = F) +
  scale_color_manual(values = continent_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(x = "GDP per capita",
       y = "Life Expectancy") +
  transition_states(continent,
                    transition_length = 5,
                    state_length = 1)

```

## Enter and Exit

```{r}

anim <- ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point(aes(color = Species), size = 2) +
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

anim +
  enter_fade() +
  exit_shrink()

```

```{r}

anim + 
  enter_fade() + enter_drift(x_mod = -1) +
  exit_shrink() + exit_drift(x_mod = 5)
 
```



```{r}

q +
  transition_states(continent,
                    transition_length = 2,
                    state_length = 1) +
  enter_fade() +
  exit_shrink()

```

## Rendering

```{r}

animate(
  anim + enter_fade() +exit_fly(y_loc = 1),
  render = av_renderer(fil = "/Users/yifeiliu/Documents/R/output/EDA/package_walkthrough/test.gif")
)

```


```{r}
animate(
  anim + enter_fade() +exit_fly(y_loc = 1) +
    exit_fade(), width = 400, height = 600, res = 35,
  render = av_renderer(fil = "/Users/yifeiliu/Documents/R/output/EDA/package_walkthrough/test.gif")
)

```



https://github.com/thomasp85/gganimate/issues/279
https://d4tagirl.com/2017/05/how-to-plot-animated-maps-with-gganimate
https://www.aliesdataspace.com/2019/05/animation-station/
https://www.blog.cultureofinsight.com/2017/09/animated-choropleth-maps-in-r/
https://github.com/thomasp85/transformr























