---
title: "ggplot themes"
author: "Yifei Liu"
date: "9/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(viridis)
library(ggthemes)
library(gcookbook)
library(hrbrthemes) # install.packages("hrbrthemes", repos = "https://cinc.rud.is")
library(ggeconodist)
packageVersion("hrbrthemes")


```

## hrbrthemes themes

This pacakge is a very focused pacakge that provides typography-centeric themes and theme components for ggplot2. It is an extract/riff of brhimisc created by request. 

### Base theme (Arial Narrow)

```{r}
ggplot(mtcars, aes(mpg, wt)) +
  geom_point() +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 scatterplot example",
       subtitle="A plot that is only useful for demonstration purposes",
       caption="Brought to you by the letter 'g'") + 
  theme_ipsum()


```

### Roboto Condensed

```{r}
ggplot(mtcars, aes(mpg, wt)) +
  geom_point() +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 scatterplot example",
       subtitle="A plot that is only useful for demonstration purposes",
       caption="Brought to you by the letter 'g'") + 
  theme_ipsum_rc()
```


### New FT Theme!
```{r}

ggplot(mtcars, aes(mpg, wt)) +
  geom_point(color = ft_cols$yellow) +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 scatterplot example",
       subtitle="A plot that is only useful for demonstration purposes",
       caption="Brought to you by the letter 'g'") + 
  theme_ft_rc()

```

### IBM Plex Sans

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_jitter(aes(color=class, fill=class), size=3, shape=21, alpha=1/2) +
  scale_x_continuous(expand=c(0,0), limits=c(1, 8), breaks=1:8) +
  scale_y_continuous(expand=c(0,0), limits=c(10, 50)) +
  scale_color_ipsum() +
  scale_fill_ipsum() +
  facet_wrap(~class, scales="free") +
  labs(
    title="IBM Plex Sans Test",
    subtitle="This is a subtitle to see the how it looks in IBM Plex Sans",
    caption="Source: hrbrthemes & IBM"
  ) +
  theme_ipsum_ps(grid="XY", axis="xy") +
  theme(legend.position="none") -> gg

flush_ticks(gg)
## theme(axis.text.x=element_text(hjust=c(0, rep(0.5, 6), 1))) +
## theme(axis.text.y=element_text(vjust=c(0, rep(0.5, 3), 1)))
```

### Titillium Web

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_jitter(aes(color=class, fill=class), size=3, shape=21, alpha=1/2) +
  scale_x_continuous(expand=c(0,0), limits=c(1, 8), breaks=1:8) +
  scale_y_continuous(expand=c(0,0), limits=c(10, 50)) +
  scale_color_ipsum() +
  scale_fill_ipsum() +
  facet_wrap(~class, scales="free") +
  labs(
    title="Titillium Web",
    subtitle="This is a subtitle to see the how it looks in Titillium Web",
    caption="Source: hrbrthemes & Google"
  ) +
  theme_ipsum_tw(grid="XY", axis="xy") +
  theme(legend.position="none") -> gg

flush_ticks(gg)
```

### Scales (Color/Fill)

```{r}

ggplot(mtcars, aes(mpg, wt)) +
  geom_point(aes(color=factor(carb))) +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 scatterplot example",
       subtitle="A plot that is only useful for demonstration purposes",
       caption="Brought to you by the letter 'g'") + 
  scale_color_ipsum() +
  theme_ipsum_rc()

```

### Scales (Axis)

```{r}
count(mpg, class) %>% 
  mutate(pct=n/sum(n)) %>% 
  ggplot(aes(class, pct)) +
  geom_col() +
  scale_y_percent() +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 column chart example with percents",
       subtitle="A plot that is only useful for demonstration purposes",
       caption="Brought to you by the letter 'g'") + 
  theme_ipsum(grid="Y") # only show certain grid


```

```{r}
ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + 
  geom_area() +
  scale_fill_ipsum() +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_comma() +
  labs(title="Age distribution of population in the U.S., 1900-2002",
       subtitle="Example data from the R Graphics Cookbook",
       caption="Source: R Graphics Cookbook") +
  theme_ipsum_rc(grid="XY") +
  theme(axis.text.x=element_text(hjust=c(0, 0.5, 0.5, 0.5, 1))) +
  theme(legend.position="bottom")
```

```{r}
update_geom_font_defaults(font_rc_light)

count(mpg, class) %>% 
  mutate(n=n*2000) %>% 
  arrange(n) %>% 
  mutate(class=factor(class, levels=class)) %>% 
  ggplot(aes(class, n)) +
  geom_col() +
  geom_text(aes(label=scales::comma(n)), hjust=0, nudge_y=2000) +
  scale_y_comma(limits=c(0,150000)) +
  coord_flip() +
  labs(x="Fuel efficiency (mpg)", y="Weight (tons)",
       title="Seminal ggplot2 column chart example with commas",
       subtitle="A plot that is only useful for demonstration purposes, esp since you'd never\nreally want direct labels and axis labels",
       caption="Brought to you by the letter 'g'") + 
  theme_ipsum_rc(grid="X")
```

### Spellcheck ggplot2 labels

```{r}
df <- data.frame(x=c(20, 25, 30), y=c(4, 4, 4), txt=c("One", "Two", "Three"))

ggplot(mtcars, aes(mpg, wt)) +
  geom_point() +
  labs(x="This is some txt", y="This is more text",
       title="Thisy is a titlle",
       subtitle="This is a subtitley",
       caption="This is a captien") +
  theme_ipsum_rc(grid="XY") -> gg

gg_check(gg)
```

Reference: [hrbrthemes : Additional Themes and Theme Components for ‘ggplot2’](https://cinc.rud.is/web/packages/hrbrthemes/index.html)


## Economist type theme

```{r}

gg <- ggplot(mammogram_costs, aes(x = city)) +
  geom_econodist(
    aes(ymin = tenth, median = median, ymax = ninetieth),
    stat = "identity", show.legend = TRUE
  ) +
  scale_y_continuous(expand = c(0,0), position = "right", limits = range(0, 800)) +
  coord_flip() +
  labs(
    x = NULL, y = NULL,
    title = "Mammoscams",
    subtitle = "United States, prices for a mammogram*\nBy metro area, 2016, $",
    caption = "*For three large insurance companies\nSource: Health Care Cost Institute"
  ) +
  theme_econodist() 

grid.newpage()

left_align(gg, c("subtitle", "title", "caption")) %>% 
  add_econodist_legend(econodist_legend_grob(), below = "subtitle") %>% 
  grid.draw()

```














