---
title: "gt walkthrough"
author: "Yifei Liu"
date: "3/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(gt) # remotes::install_github("rstudio/gt")
library(emo) # devtools::install_github("hadley/emo")
detach("package:dplyr", unload = TRUE)
library(dplyr)
```

## introducing the gt packages

```{r}
exibble

exibble %>% gt()

tab <- exibble %>% gt(
  rowname_col = "row",
  groupname_col = "group"
)

tab_2 <- tab %>%
  fmt_scientific(
    columns = vars(num),
    decimals = 3
  )

tab_2

tab_3 <- tab_2 %>%
  fmt_date(
    columns = vars(date),
    rows = grepl("^[a-d]", char),
    date_style = 6
  )

tab_3

```


```{r}

info_date_style()


```


```{r}
# Anyway, let's get ride of some columns
tab_4 <- tab_3 %>%
  cols_hide(
    columns = vars(char, fctr, time, datetime)
  )

tab_4

#Format the currency columns to have

tab_5 <- tab_4 %>%
  fmt_currency(columns = vars(currency),
               currency = "GBP")

tab_5

# Add conditional footnotes to some of the 'currency' values

tab_6 <- tab_5 %>%
  tab_footnote(
    footnote = "These are lower prices.",
    locations = cells_body(
      columns = vars(currency),
      rows = currency < 20
    )
  )

tab_6

# Add a footntos to the `currency` column label itself

tab_7 <- tab_6 %>%
  tab_footnote(
    footnote = "All values are in GBP",
    locations = cells_column_labels(
      columns = vars(currency)
    )
  )

tab_7

# Add a header to the table (with a title and a subtitle, use md for markdown format)

tab_8 <- tab_7 %>%
  tab_header(
    title = "An Exibble Example",
    subtitle = md("Uses the *exibble* dataset from **gt**")
  )
tab_8



```


pizza_tab_email dataset



```{r}

sizes_order <- c("S", "M", "L", "XL", "XXL")
types_order <- c("Classic", "Chicken", "Supreme", "Veggie")

pizzaplace %>% 
  mutate(type = str_to_title(type),
         size = factor(size, levels = sizes_order),
         type = factor(type, levels = types_order)) %>%
  group_by(type, size) %>%
  summarize(
    pies = n(),
    income = sum(price)
  ) %>%
  arrange(type, size) %>%
  gt(rowname_col = "size") %>%
  fmt_currency(
    columns = vars(income),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(pies),
    use_seps = TRUE,
    decimals = 0
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = "pies",
    fns = list(TOTAL = "sum"),
    formatter = fmt_number,
    use_seps = TRUE,
    decimals = 0
    ) %>%
   summary_rows(
    groups = TRUE,
    columns = "income",
    fns = list(TOTAL = "sum"),
    formatter = fmt_currency,
    currency = "USD"
    ) %>%
  tab_options(
    summary_row.background.color = "#FFFEEE",
    table.font.size = px(14),
    row_group.background.color = "#E6EFFC",
    row_group.padding = "5px") %>%
  cols_label(
    pies = "Pizzas",
    income = "Income"
    ) %>%
  tab_header(
    title = paste0("My ", emo::ji("pizza"), " sales in 2015"),
    subtitle = "Split by the type of pizza and the size"
    ) %>%
  tab_footnote(
    footnote = md("Only **The Greek Pizza** comes in this size."),
    locations = cells_stub(rows = 4:5)
    ) %>%
  tab_footnote(
    footnote = "The small-sized classic pizzas sold the most.",
    locations = cells_body(columns = 1, rows = 1)
    ) %>%
  as_raw_html()
 
  



```


## Introduction to Creating gt Tables

```{r}
# take the biggest islands in the world from `islands` dataset
islands_tbl <- 
  dplyr::tibble(
    name = names(islands),
    size = islands
  ) %>%
  dplyr::arrange(desc(size)) %>%
  dplyr::slice(1:10)

islands_tbl

```



```{r}
# create a display table showing ten of the largest islands in the world
gt_tbl <- gt(data = islands_tbl)

gt_tbl



```

```{r}

gt_tbl_1 <- gt_tbl %>%
  tab_header(
    title = "Large Landmasses of the World",
    subtitle = "The Top Ten Largest are presented"
  )

gt_tbl_1

```

```{r}

gt(islands_tbl[1:2, ]) %>%
  tab_header(
    title = md("**Large Landmess of the World**"),
    subtitle = md("the *top two* larest are presented")
  )


```

```{r}
gt_tbl_2 <- gt_tbl_1 %>%
  tab_source_note(
    source_note = "Source: The World Almanac and Book of Facts, 1875, page 406."
  ) %>%
  tab_source_note(
    source_note = md("Refreence: McNeil, D.R. (1977) *Interactive Data ANalysis*. Wiley")
  )

gt_tbl_2

```

```{r}

gt_tbl_3 <- gt_tbl_2 %>%
  tab_footnote(
    footnote = "The Americans.",
    locations = cells_body(
      columns = vars(name),
      rows = 3:4
    )
  )

gt_tbl_3

```

```{r}

# Determine the row that contains the largest landmass ('Asia')

largest <- islands_tbl %>%
  slice(1) %>%
  pull(name)

# create two additional footnotes, using the `columns` and `where` arguments of `data_cells()`

gt_tbl_4 <- gt_tbl_3 %>%
  tab_footnote(
    footnote = md("The **largest** by area."),
    locations = cells_body(
      columns = vars(size),
      rows = name == largest
    )
  ) %>%
  tab_footnote(
    footnote = "The lowest by population.",
    locations = cells_body(
      columns = vars(size),
      rows = size == min(size)
    )
  )

gt_tbl_4

```

The stub

```{r}

gt_tbl <- islands_tbl %>%
  gt(rowname_col = "name")

gt_tbl
```

notice the landmass names are off the left in an unstriped area? That's the **stub**. 
```{r}

gt_tbl_1 <- gt_tbl %>%
  tab_stubhead(label = "landmass")

gt_tbl_1
```

```{r}

gt_tbl_2 <- gt_tbl_1 %>%
  tab_header(
    title = "Largest Landmass of the World",
    subtitle = "The top ten largest are presented"
  ) %>%
  tab_source_note(
    source_note = "Source: The World Almanac and Book of Facts, 1975, page 406."
  ) %>%
  tab_source_note(
    source_note = md("Reference: McNeil,  D. R. (1977) *Interactive Data Analysis*. Wiley.")
  ) %>%
  tab_footnote(
    footnote = md("The **largest** by area"),
    locations = cells_body(
      columns = vars(size),
      rows = size == max(size)
    )
  ) %>%
  tab_footnote(
    footnote = "The lowest by population.",
    locations = cells_body(
      columns = vars(size),
      rows = contains("arc")
    )
  )

gt_tbl_2

```

let's divided rows into groups, creating row groups. and results in a display of _row group labels_ right above the each group. 

```{r}

gt_tbl_3 <- gt_tbl_2 %>%
  tab_row_group(
    group = "continent",
    rows = 1:6
  ) %>%
  tab_row_group(
    group = "country",
    rows = c("Australia", "Greenland")
  ) %>%
  tab_row_group(
    group = "subcontinent",
    rows = c("New Guinea", "Borneo")
  )

gt_tbl_3

```

The column labels


```{r}

airquality_m <- airquality %>%
  mutate(Year = 1973L) %>%
  slice(1:10)

# create a display table using the `airquality` dataset; arrange columns into groups

gt_tbl <-gt(airquality_m) %>%
  tab_header(
    title = "New York Air Quality Measurements",
    subtitle = "Daily measurements in New York City (May 1-10, 1973)"
  ) %>%
  tab_spanner(
    label = "Time",
    columns = vars(Year, Month, Day)
  ) %>%
  tab_spanner(
    label = "Measurement",
    columns = vars(Ozone, Solar.R, Wind, Temp)
  )

gt_tbl

```

we need to do two more things to make this table presentable

* Move the time columns to the beginning of the series (use cols_move_to_start)
* Customize the column labels so that they are more descriptive (use cols_label)

```{r}

gt_tbl_1 <- gt_tbl %>%
  cols_move_to_start(
    columns = vars(Year, Month, Day)
  ) %>%
  cols_label(
    Ozone = html("Ozone,<br>ppdv"),
    Solar.R = html("Solar R.,<br>cal/m<sup>2</sup>"),
    Wind = html("Wind,<b>mph"),
    Temp = html("Temp,<br>&deg;F")
  )

gt_tbl_1

```



## gt Datasets

1. countrypops 

```{r}

# Get vectors of 2-letter country codes for
# each region of Oceania
Australasia <- c("AU", "NZ")
Melanesia <- c("NC", "PG", "SB", "VU")
Micronesia <- c("FM", "GU", "KI", "MH", "MP", "NR", "PW")
Polynesia <- c("PF", "WS", "TO", "TV")

# Create a gt table based on a preprocessed `countrypops`
countrypops_dbl <- countrypops %>%
  dplyr::filter(country_code_2 %in% c(
    Australasia, Melanesia, Micronesia, Polynesia)
  ) %>%
  dplyr::filter(year %in% c(1995, 2005, 2015)) %>%
  dplyr::mutate(region = case_when(
    country_code_2 %in% Australasia ~ "Australasia",
    country_code_2 %in% Melanesia ~ "Melanesia",
    country_code_2 %in% Micronesia ~ "Micronesia",
    country_code_2 %in% Polynesia ~ "Polynesia",
  )) %>%
  tidyr::spread(key = year, value = population) %>%
  dplyr::arrange(region, desc(`2015`)) %>%
  dplyr::select(-starts_with("country_code")) 

gt_tbl <- countrypops_dbl %>%
  gt(rowname_col = "country_name",
     groupname_col = "region")

gt_tbl %>%
  tab_header("Population of Oceania's Countries in 1995, 2005, and 2015") %>%
  tab_spanner(
    label = "Total population",
    columns = vars(`1995`, `2005`, `2015`)
  ) %>%
  fmt_number(
    columns = vars(`1995`, `2005`, `2015`),
    decimals = 0, 
    use_seps = TRUE
  )

```

2. sza

```{r}

sza_tbl <- sza %>%
  filter(latitude == 20,
         !is.na(sza)) %>%
  select(-latitude) %>%
  spread(key = "tst", value = sza)

gt_tbl <- sza_tbl %>%
  gt(rowname_col = "month") %>%
  fmt_missing(
    columns = TRUE,
    missing_text = ""
  ) %>%
  tab_stubhead(label = html("month<br>(20&deg;N)")) %>%
  tab_header(title = html("&#x2600; Solar Zenith Angles &#x2600;")) %>%
  tab_options(
    column_labels.font.size = "smaller",
    table.font.size = "smaller",
    data_row.padding = px(3)
  )

gt_tbl

```

3. gt cars

```{r}
# create a gt table based on a preprocessed `gtcars`

gt_tbl <- gtcars %>%
  dplyr::filter(ctry_origin == "Germany") %>%
  dplyr::group_by(mfr) %>%
  dplyr::top_n(2, msrp) %>%
  dplyr::ungroup() %>%
  dplyr::select(mfr, model, drivetrain, msrp) %>%
  gt()

gt_tbl %>%
  tab_header(title = "Select German Automobiles") %>%
  cols_merge(
    columns = vars(mfr, model),
    hide_columns = vars(model)
  ) %>%
  text_transform(
    locations = cells_body(columns = vars(drivetrain)), 
    fn = function(x) toupper(x)
  ) %>%
  fmt_currency(
    columns = vars(msrp),
    currency = "USD",
    decimals = 0
  ) %>%
  cols_label(
    mfr = "Car",
    drivetrain = "Derivetrain",
    msrp = "MSRP"
  ) %>%
  tab_footnote(
    footnote = "Prices in USD",
    locations = cells_column_labels(vars(msrp))
  ) %>%
  tab_footnote(
    footnote = "AWD = All Wheel Drive, RWD = Real Wheel Drive.",
    locations = cells_column_labels(vars(drivetrain))
  ) %>%
  opt_footnote_marks(marks = "letters")
  
```

4. An sp500 example

```{r}
# Define the start and end dates for the data range
start_date <- "2010-06-02"
end_date <- "2010-06-15"

# The HTML decimal references for the black
# up- and down-pointing triangles are: #9650 and #9660;
# use an in-line style to apply color
up_arrow <- "<span style=\"color:green\">&#9650;</span>"
down_arrow <- "<span style=\"color:red\">&#9660;</span>"

gt_dbl <- sp500 %>%
  dplyr::filter(date >= start_date & date <= end_date) %>%
  dplyr::select(-adj_close) %>%
  gt()

gt_dbl %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue::glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 7
  ) %>% 
  fmt_currency(
    columns = vars(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(volume),
    scale_by = 1 / 1E9,
    pattern = "{x}B"
  ) %>%
  text_transform(
    locations = cells_body(
      columns = "close",
      rows = open < close
    ),
    fn = function(x) paste(x, up_arrow)
  ) %>%
  text_transform(
    locations = cells_body(
      columns = "close",
      rows = open > close
    ),
    fn = function(x) paste(x, down_arrow)
  ) %>%
  cols_label(
    date = "Date",
    open = "Open",
    high = "High",
    low = "Low",
    close = "Close",
    volume = "Volume"
  )


```

5. a pizzaplace example

```{r}

# create a gt table based on a preprocessed `pizzaplace`

gt_tbl <- pizzaplace %>%
  group_by(type, size) %>%
  summarize(revenue = sum(price),
            sold = n()) %>%
  gt(rowname_col = "size")

gt_tbl %>%
  tab_header("Pizza Sold in 2015") %>%
  fmt_currency(
    columns = vars(revenue),
    currency = "USD",
    decimals = 0
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(sold),
    fns = list(Total = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(revenue),
    fns = list(Total = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  tab_options(
    summary_row.background.color = "#ACEACE",
    row_group.background.color = "#FFEFDB"
  )
  

```


6. An exibble Example

```{r}

gt_tbl <- exibble %>%
  gt(
    rowname_col = "row",
    groupname_col = "group"
  )

gt_tbl %>%
  fmt_number(
    columns = vars(num),
    decimals = 2
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 6
  ) %>% 
  fmt_currency(
    columns = vars(currency),
    currency = "EUR"
  ) %>%
  tab_options(
    column_labels.font.size = "small",
    table.font.size = "small",
    row_group.font.size = "small",
    data_row.padding = px(3)
  )

```

### Case study: gtcars

```{r}

dplyr::glimpse(gtcars)

```
reduce 47 example to only 8

```{r}

gtcars_8 <- gtcars %>%
  group_by(ctry_origin) %>%
  top_n(2) %>%
  ungroup() %>%
  filter(ctry_origin != "United Kingdom")

glimpse(gtcars_8)

```

display table from this dataset, discard and highlight some info

1. putting the cars into characteristic groups, by the car manufacture;s country of origin

```{r}
# to arrange observation group together by country of origin, we need to use group_by options
gtcars_8 %>%
  group_by(ctry_origin) %>%
  gt()

# getting the row groups into prefered order by using arrange function

gtcars_8 %>%
  group_by(ctry_origin) %>%
  arrange(mfr, desc(msrp)) %>%
  gt()

# use factor leves to get a more particular ordering with arrange. first we change order of country by our own preference and then arrange by mfr in a descend order
order_countries <- c("Germany", "Italy", "United States", "Japan")

gtcars_8 %>%
  arrange(factor(ctry_origin, levels = order_countries), mfr, desc(msrp)) %>%
  group_by(ctry_origin) %>%
  gt()


# reorder the table rows by our specific ordering of groups

tab <- gtcars_8 %>%
  arrange(factor(ctry_origin, levels = order_countries), mfr, desc(msrp)) %>%
  mutate(car = paste(mfr, model)) %>%
  select(-mfr, -model) %>%
  group_by(ctry_origin) %>%
  gt(rowname_col = "car")

tab

```

Hiding and moving some columns

why not just remove this columns, in some case you will need variables for conditional statements within **gt** but won't want to display them in the end.

```{r}

# Use a few `cols_*()` functions to hide and move columns

tab_1 <- tab %>%
  cols_hide(columns = vars(drivetrain, bdy_style)) %>%
  cols_move(
    columns = vars(trsmn, mpg_c, mpg_h),
    after = vars(trim)
  )

tab_1

```

putting columns into groups

sometimes useful to arrange variables/columns into groups by using spanner column labels.

```{r}
# put the first three columsn under a spanner column with the label 'Performance'
tab_2 <- tab_1 %>%
  tab_spanner(label = "Performance",
              columns = vars(mpg_c, mpg_h, hp, hp_rpm, trq, trq_rpm))


tab_2
```

Merging columns together and labeling them

```{r}
# perform three column merges to better present MPG, HP, and torque; relabel all the remaining columns for a nicer-looking presentation

tab_3 <- tab_2  %>%
  cols_merge(
    vars(mpg_c, mpg_h),
    hide_columns = vars(mpg_h),
    pattern = "{1}c<br>{2}h"
    ) %>%
  cols_merge(
    vars(hp, hp_rpm),
    hide_columns = vars(hp_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) %>%
  cols_merge(
    vars(trq, trq_rpm),
    hide_columns = vars(trq_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) %>%
  cols_label(
    mpg_c = "MPG",
    hp = "HP",
    trq = "Torque",
    year = "Year",
    trim = "Trim",
    trsmn = "Transmission",
    msrp = "MSRP"
  )

tab_3

```

Using formatter functions

```{r}

# format the msrp column to USD currency

tab_4 <- tab_3 %>%
  fmt_currency(columns = vars(msrp),
               currency = "USD",
               decimals = 0)


tab_4
```

column Alignment and style chagnes

```{r}

# Center-align three columns in the gt table and modify the text size of a few columns of data

tab_5 <- tab_4 %>%
  cols_align(
    align = "center",
    columns = vars(mpg_c, hp, trq)
  ) %>%
  tab_style(
    style = cell_text(size = px(12)),
    locations = cells_body(
      columns = vars(trim, trsmn, mpg_c, hp, trq))
  )

tab_5

```


Text transforms

```{r}

tab_6 <- tab_5 %>%
  text_transform(
    locations = cells_body(columns = vars(trsmn)),
    fn = function(x) {

      # The first character of `x` always
      # indicates the number of transmission speeds
      speed <- substr(x, 1, 1)

      # We can carefully determine which transmission
      # type we have in `x` with a `dplyr::case_when()`
      # statement
      type <-
        dplyr::case_when(
          substr(x, 2, 3) == "am" ~ "Automatic/Manual",
          substr(x, 2, 2) == "m" ~ "Manual",
          substr(x, 2, 2) == "a" ~ "Automatic",
          substr(x, 2, 3) == "dd" ~ "Direct Drive"
        )

      # Let's paste together the `speed` and `type`
      # vectors to create HTML text replacing `x`
      paste(speed, " Speed<br><em>", type, "</em>")
    }
  )

tab_6

```

table header: title and subtitle

```{r}

tab_7 <- tab_6 %>%
  tab_header(
    title = md("The Cars of **gtcars**"),
    subtitle ="There are some fine automobiles"
  )

tab_7

```

Adding a source citation

```{r}

tab_8 <- tab_7 %>%
  tab_source_note(
    source_note = md("Source: Various pages within the Edmonds Website")
  )

tab_8

```

Using complete grcars table and adding footnotes

1. Identifying the car with the best gas milage (city)
2. Identifying the car with the highest horsepower
3. Stating the currency of the MSRP

```{r}
best_gas_milage_city <- gtcars %>%
  arrange(desc(mpg_c)) %>%
  slice(1) %>%
  mutate(car = paste(mfr, model)) %>%
  pull(car)

# use dplyr function to get the car witht he highest horsepower this will be used to target the correct cell for a footnote

highest_horsepower <- gtcars %>%
  arrange(desc(hp)) %>%
  slice(1) %>%
  mutate(car = paste(mfr, model)) %>%
  pull(car)

# create a display table with gtcars, using all of the previous statement piped together + addiitonal 'tab_footnotes' stmts

tab <- gtcars %>%
  arrange(factor(ctry_origin, levels = order_countries),
          mfr, desc(msrp)) %>%
  mutate(car = paste(mfr, model)) %>%
  select(-mfr, -model) %>%
  group_by(ctry_origin) %>%
  gt(rowname_col = "car") 


tab %>%
  cols_hide(columns = vars(drivetrain, bdy_style)) %>%
  cols_move(
    columns = vars(trsmn, mpg_c, mpg_h),
    after = vars(trim)
  ) %>%
  tab_spanner(
    label = "Performance",
    columns = vars(mpg_c, mpg_h, hp, hp_rpm, trq, trq_rpm)
  ) %>%
  cols_merge(
    vars(mpg_c, mpg_h),
    hide_columns = vars(mpg_h),
    pattern = "{1}c<br>{2}h"
  ) %>%
  cols_merge(
    vars(hp, hp_rpm),
    hide_columns = vars(hp_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) %>%
  cols_merge(
    vars(trq, trq_rpm),
    hide_columns = vars(trq_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) %>%
   cols_label(
    mpg_c = "MPG",
    hp = "HP",
    trq = "Torque",
    year = "Year",
    trim = "Trim",
    trsmn = "Transmission",
    msrp = "MSRP"
  ) %>%
  fmt_currency(
    columns = vars(msrp),
    currency = "USD",
    decimals = 0
  ) %>%
  cols_align(
    align = "center",
    columns = vars(mpg_c, hp, trq)
  ) %>%
  tab_style(
    style = cell_text(size = px(12)),
    locations = cells_body(
      columns = vars(trim, trsmn, mpg_c, hp, trq)
    )
  ) %>%
  text_transform(
    locations = cells_body(columns = vars(trsmn)),
    fn = function(x) {

      speed <- substr(x, 1, 1)

      type <-
        dplyr::case_when(
          substr(x, 2, 3) == "am" ~ "Automatic/Manual",
          substr(x, 2, 2) == "m" ~ "Manual",
          substr(x, 2, 2) == "a" ~ "Automatic",
          substr(x, 2, 3) == "dd" ~ "Direct Drive"
        )

      paste(speed, " Speed<br><em>", type, "</em>")
    }
  ) %>%
  tab_header(
    title = md("The Cars of **gtcars**"),
    subtitle = "These are some fine automobiles"
  ) %>%
  tab_source_note(
    source_note = md(
      "Source: Various pages within the Edmonds website.")
  ) %>%
  tab_footnote(
    footnote = md("Best gas mileage (city) of all the **gtcars**"),
    locations = cells_body(
      columns = vars(mpg_c),
      rows = best_gas_milage_city
    )
  ) %>%
    tab_footnote(
    footnote = md("The highest horsepower of all the **gtcars**."),
    locations = cells_body(
      columns = vars(hp),
      rows = highest_horsepower)
  ) %>%
  tab_footnote(
    footnote = "All prices in U.S. dollars (USD).",
    locations = cells_column_labels(columns = vars(msrp))
  )



```




https://gt.rstudio.com/index.html




