---
title: "pdftools walkthrough"
author: "Yifei Liu"
date: "2/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(pdftools)
library(tidyverse)
library(glue)
theme_set(theme_minimal())

```

# Example 1: Basic

[Extracting data from a PDF into R](https://rstudio-pubs-static.s3.amazonaws.com/415060_553527fd13ed4f30aae0f1e4483aa970.html)


1. create a new file
2. Load packages
3. load pdf
```{r}
pdf <- pdf_text("/Users/yifeiliu/Documents/R/data/EDA/package_walkthrough/oregon_grass_and_legume_seed_crops_preliminary_estimates_2017.pdf") %>%
  readr::read_lines()

pdf


```


```{r}
(file_data_raw <- pdf[-c(1:3, 6:8, 20:35)])

(file_data_raw[3:13] %>%
    str_squish() %>%
    strsplit(split = " ") -> file_data)


var_lines <- c("Species", "Arceage", "Yield", "Production", "Price", "Value")

file_data[[6]] <- c("Orchard", "grass", "15,190","1,046","15,889","225.00","35,750")

(df <- plyr::ldply(file_data))

df %>%
  unite(V1.2, V1, V2, sep = "_") %>%
  setNames(var_lines) %>%
  mutate_at(var_lines[2:6], parse_number)

```

summary: very basic example, does't involve a lot of regular expression and not very complicated. Don't involve issues such as contain missing value, (make str_squish difficult), one value occupied more than one row, (such as long country name, when use read_lines, it split one entry into three sometimes.), etc


# Example 2: Advance


Download files

```{r}
library(ggthemes)
country <- c("lux", "fra", "deu", "usa", "prt", "gbr")

url <- "http://www.who.int/diabetes/country-profiles/{country}_en.pdf?ua=1"

```

```{r}

(urls <- glue(url))

pdf_names <- glue("/Users/yifeiliu/Documents/R/data/EDA/package_walkthrough/report_{country}.pdf")

walk2(urls, pdf_names, download.file, mode = "wb")


```

read files

```{r}

raw_text <- map(pdf_names, pdf_text)

raw_text[[1]]

```

```{r}

clean_table <- 

```

```{r}
x_1 <- str_split(raw_text[[1]], "\n", simplify = T)
x_2 <- x_1[1, 1] %>%
  stringr::str_squish() %>%
  stringr::str_extract(".+?(?=\\sTotal)")

x_1[1, 1] %>%
  stringr::str_squish() %>%
  stringr::str_extract(".+?(?=\\sTotal)")
```








