---
title: "Tidyquant walkthrough"
author: "Yifei Liu"
date: "12/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load required library

```{r, message=F}
library(tidyquant)
library(knitr)
library(kableExtra)
library(scales)
library(finstr)
library(XBRL)
library(tidyverse)
library(scales)
detach("package:dplyr", unload=TRUE)
library(dplyr)
theme_set(theme_minimal())
```


We can get the stock price data of FAANG and XLK

```{r}
asset_price  <- c("AAPL", "GOOGL","AMZN", "NFLX", "FB") %>%
  tq_get(get = "stock.prices",
         from = "2010-01-01",
         to = "2018-12-1")  

Ra <- asset_price %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "Ra")

Rb <-  tq_get("XLK", get = "stock.prices",
         from = "2010-01-01",
         to = "2018-12-1") %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "Rb")

Ra
Rb
RaRb <- left_join(Ra, Rb, by = "date")
```

Plot the Ra data

```{r}
asset_price %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  theme_tq() +
  labs(x = "", y = "") +
  scale_y_continuous(labels = dollar_format())

```



Now we can quick do some performance analysis on the stocks.
1. CAPM
2. Information Ratio
3. VaR, ES
4. table.Correlation

```{r}

RaRb %>%
  tq_performance(Ra,
                 Rb,
                 performance_fun = table.CAPM,
                 scale = 12) %>%
  select(Alpha, Beta)

RaRb %>%
  tq_performance(Ra,
                Rb,
                performance_fun = InformationRatio,
                scale = 12)
Ra %>%
  tq_performance(performance_fun = VaR,
                 method = "historical")

RaRb %>%
  tq_performance(Ra = Ra, Rb = Rb, 
                 performance_fun = table.Correlation)

RaRb %>%
  tq_performance(Ra = Ra, 
                 performance_fun = AverageDrawdown)


```




## Portfolios


Efficient Frontier

```{r}

RaRb %>%
  tq_performance(Ra = Ra, 
                 performance_fun = table.AnnualizedReturns) %>%
  ggplot(aes(AnnualizedStdDev, AnnualizedReturn)) +
  geom_point() +
  geom_label(aes(label = symbol, color = symbol), show.legend = F) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Annualized sd", y = "Annualized Return") +
  theme_tq()


```


We can use weight vector such as rep(0.2, 5) 

```{r}
wts <- rep(0.2, 5)

portfolio_returns <- Ra %>%
  tq_portfolio(assets_col = symbol,
               returns_col = Ra,
               weights = wts,
               col_rename = "Ra")

portfolio_returns

```

We can aggregated the portfolio return and benchmark returns together

```{r}
RaRb_portfolio <- portfolio_returns %>%
  left_join(Rb, by = "date")


```


We can perofrmance different type of perofrmance analysis on this dataset

```{r}
RaRb_portfolio %>%
  tq_performance(Ra,
                 Rb,
                 performance_fun = table.CAPM) 

```

We can now try to create multipel portfolio use tq_repeat_df

```{r}
Ra_multi <- Ra %>%
  ungroup(symbol) %>%
  tq_repeat_df(n = 4)

weigths <- c(rep(0.2, 5),
             rep(0.25, 4), 0,
             rep(0.1, 3), 0.35, 0.35,
             rep(0.15, 4), 0.4)

weights_table <- tibble(c("AAPL", "GOOGL","AMZN", "NFLX", "FB")) %>%
  tq_repeat_df(n = 4) %>%
  bind_cols(tibble(weigths)) %>%
  group_by(portfolio) %>%
  setNames(c("portfolio",  "stocks", "weights"))

portfolio_multi_returns <- Ra_multi %>%
  tq_portfolio(assets_col = symbol,
               returns_col = Ra,
               weights = weights_table,
               col_rename = "Ra")
```


After create different weight portfolio, we can see how each one perform by aggregate this data with baseline data.

```{r}
RaRb_multi_portfolio <- portfolio_multi_returns %>%
  left_join(Rb, by = "date")
```

We can calculate each portfolio permance. These is full list of function that are avaibale in tq_peroformance_fun_options

```{r}
RaRb_multi_portfolio %>%
  tq_performance(Ra,
                 Rb,
                 performance_fun = table.CAPM,
                 scale = 12) %>%
  select(portfolio, Alpha, Beta) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)

RaRb_multi_portfolio %>%
  tq_performance(Ra,
                 Rb,
                 performance_fun = InformationRatio,
                 scale = 12) %>%
  setNames(c("Portfolio", "Information Ratio")) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)

RaRb_multi_portfolio %>%
  tq_performance(Ra,
                 performance_fun = Return.annualized) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)



```

## Chart the returns

We can not visulize the return use ggplot

```{r}

portfolio_multi_returns %>%
  ggplot(aes(x = date, y = Ra)) +
  geom_bar(stat = "identity", fill = palette_light()[[1]]) +
  facet_wrap(~ portfolio) +
  labs(title = "Portfolio Return",
       x = "",
       h = "Monthly Return") +
  geom_smooth(method = "glm") +
  theme_tq() +
  scale_color_tq() +
  scale_y_continuous(labels = percent_format())

```

This is one way to show the return are, but as we all know, return rate may be a little bit difficult to interpret. We can use dollar amount to see how money investment will change. 

```{r}
Ra_multi %>%
  tq_portfolio(assets_col = symbol,
               returns_col = Ra,
               weights = weights_table,
               col_rename = "Ra",
               wealth.index = T) %>%
  mutate(investment_grwoth = 10000 * Ra) %>%
  ggplot(aes(x = date, y = investment_grwoth, color = factor(portfolio))) +
  geom_line(size = 2) +
  labs(title = "Portfolio Return",
       x = "",
       h = "Monthly Return") +
  geom_smooth(method = "loess") +
  theme_tq() +
  scale_color_tq() +
  scale_y_continuous(labels = dollar)

```


## key ratio 

For any given stocks, the histrocial key ratios are avaibale for 10 years. 

```{r}
aapl_key_ratios <- tq_get("AAPL", get = "key.ratios")

```

```{r}
aapl_key_ratios %>%
  dplyr::filter(section == "Valuation Ratios") %>%
  unnest() %>% 
  ggplot(aes(date, value)) +
  geom_line(aes(color = factor(category)) ,size = 1) +
  theme_tq() +
  scale_color_tq()


aapl_key_ratios %>%
  dplyr::filter(section == "Valuation Ratios") %>%
  unnest()
```

Base on these example we can try to compare different firm financial ratio

```{r}
stocks <- c("AAPL", "GOOGL","AMZN", "NFLX", "FB")

faang_key_ratio <- tq_get(stocks, get = "key.ratios")

```

Have a look at different firlm valutaion ratio

```{r}
faang_key_ratio %>%
  dplyr::filter(section == "Valuation Ratios") %>%
  unnest() %>%
  ggplot(aes(date, value)) +
  geom_line(aes(color = factor(symbol)),
            size = 1) +
  facet_wrap(~ factor(category), scales = "free_y") +
  theme_tq() +
  scale_color_tq()
```


### Small Cap

```{r}
russell_2000 <- tq_index("RUSSELL2000") %>%
  tbl_df()
  
rusell_2000_price <- tq_get(x = "^RUT")

```

```{r}
russell_2000 %>%
  count(sector, wt = weight) %>%
  mutate(sector = fct_reorder(sector, n)) %>%
  ggplot(aes(sector, n)) +
  geom_col(fill = "midnightblue") +
  coord_flip() +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "", y = "")

```



























