---
title: "timetk_walkthrough"
author: "Yifei Liu"
date: "8/5/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# timetk

## plotting time series

```{r}

library(tidyverse)
library(lubridate)
library(timetk)

# Setup for the plotly charts (# FALSE returns ggplots)
interactive <- TRUE

```
### plotting single time series


```{r}

taylor_30_min

```

```{r}

taylor_30_min %>%
  plot_time_series(date, value,
                   .interactive = interactive,
                   .plotly_slider = TRUE)

```

### plotting groups

```{r}

m4_daily %>%
  group_by(id) %>%
  plot_time_series(date, value,
                   .facet_ncol = 2, .facet_scales = "free",
                   .interactive = interactive)

```


### Visualizing transformation & sub-groups

```{r}

m4_hourly %>%
  group_by(id) %>%
  plot_time_series(date, log(value),
                   .color_var = week(date),
                   .facet_ncol = 2,
                   .facet_scales = "free",
                   .interactive = interactive)

```


### Static ggplot2 Visualization & Customizations

```{r}

taylor_30_min %>%
  plot_time_series(date, value,
                   .color_var = month(date, label = T),
                   .interactive = FALSE) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(title = "Taylor's MegaWatt Data",
       x = "Date (30-min intervals)",
       y = "Energy Demand (MW)",
       color = "Month")


```


## plotting Seasonality and correlation

### Correlation plots

Grouped ACF diagnostics

```{r}
m4_hourly %>%
  group_by(id) %>%
  plot_acf_diagnostics(
    date, value,
    .lags = "7 days",
    .interactive = TRUE
  )


```

### Seasonality

```{r}

taylor_30_min %>%
  plot_seasonal_diagnostics(date, value,
                            .interactive = TRUE)

```

grouped seasonal visualizations

```{r}

m4_hourly %>%
  group_by(id) %>%
  plot_seasonal_diagnostics(date, 
                            value, .interactive = TRUE)

```


STL Diagnostics

```{r}

m4_hourly %>%
  group_by(id) %>%
  plot_stl_diagnostics(
    date, value,
    .frequency = "auto", .trend = "auto",
    .feature_set = c("observed", "season", "trend", "remainder"),
    .interactive = TRUE
  )

```



## Time Series COversion

> timetk: A toolkit for time series analysis in R










