---
title: "Fama French"
author: "Yifei Liu"
date: "1/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, message=F, warning=F}
library(tidyquant)
library(tidyverse)
library(timetk)
library(tibbletime)
library(broom)
library(scales)
library(highcharter)
library(rsample)
library(yardstick)
library(awtools) #devtools::install_github('awhstin/awtools')
library(hrbrthemes)
detach("package:dplyr", unload=TRUE)
library(dplyr)
theme_set(theme_minimal())
```



## Calculated beta in capital asset pricing model

get index price and return


```{r}
symbols <- c("SPY","EFA", "IJS", "EEM","AGG")

stock_price <- tq_get(symbols,
                      get = "stock.prices",
                      from = "2012-12-31",
                      to = "2017-12-31") %>%
  group_by(symbol)



stock_return <- stock_price %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               tpe = "log",
               col_rename = "Ra") %>%
  filter(date > "2012-12-31")

weight <- c(0.25, 0.25, 0.2, 0.2, 0.1)

portfolio_re <- stock_return %>%
  tq_portfolio(assets_col = symbol,
               returns_col = Ra,
               weights = weight,
               col_rename = "Ra",
               rebalance_on = "months")
  



# get market return, in this case we use S&P 500 as the market. 

spy_re <- tq_get("SPY",
                 get = "stock.prices",
                 from = "2012-12-31",
                 to = "2017-12-31") %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               type = "log",
               col_rename = "Rm") %>%
  filter(date > "2012-12-31")

portfolio_mk_re <- portfolio_re %>%
  inner_join(spy_re, by = "date") 
  

```

We now have the monthly balance portfolio and market return. Now we can calculated the CAPM beta

$${\beta}_{portfolio} = cov(R_p, R_m)/\sigma_m $$
```{r}
cov(portfolio_mk_re$Ra, portfolio_mk_re$Rm) / var(portfolio_mk_re$Rm)


```

We can calculated Beta by calculated beta fro each of our assets. 

$${\beta}_{portfolio} ={\sum_{i=1}^n}W _i~{\beta}_i $$

```{r}
assets_beta <- stock_return %>%
  nest(- symbol)


assets_beta <- assets_beta %>%
  mutate(model = map(data, ~ lm(Ra ~ spy_re$Rm, data = .)),
         model = map(model, tidy)) %>%
  unnest(model)

assets_beta %>%
  filter(term == "spy_re$Rm")

beta_estimated <- assets_beta %>%
  filter(term == "spy_re$Rm")

cov(portfolio_mk_re$Ra, portfolio_mk_re$Rm) / var(portfolio_mk_re$Rm)

sum(beta_estimated$estimate * weight)


```

we can see the individual asset beta together equal portfolio beta. 

use tidyverse and broom to calculate beta and alpha for individual asset and beta for portfolio.

```{r}

portfolio_beta <- portfolio_mk_re %>%
  do(model = lm(Ra ~ Rm, data = .)) %>%
  tidy(model) %>%
  mutate(term = c("alpha", "beta"))


# individual asset beta and alpha
beta_tidy <- stock_return %>%
  group_by(symbol) %>%
  do(model = lm(Ra ~ spy_re$Rm, data = .)) %>%
  tidy(model) %>%
  mutate(term = c("alpha", "beta"))

# portfolio beta
beta_pf_tidy <- portfolio_re %>%
  do(model = lm(Ra ~ spy_re$Rm, data = .)) %>%
  tidy(model) %>%
  mutate(term = c("alpha", "beta"))
```
Use CAPM beta in tidyquant to calculate Beta.


```{r}
beta_tq <- portfolio_mk_re %>%
  tq_performance(Ra = Ra,
                 Rb = Rm,
                 performance_fun = CAPM.beta)
beta_tq

```

## Visualzing the CAPM 

## Visualzing the relationship between portfolio returns, risk and market returns

```{r}
stock_return %>%
  summarize(exp_re = mean(Ra),
            sd = sd(Ra)) %>%
  ggplot(aes(sd, exp_re, color = symbol)) +
  geom_label(aes(label = symbol), show.legend = F, size = 4) +
  geom_text(aes(x = sd(portfolio_re$Ra), y = mean(portfolio_re$Ra)), label = "Portfolio", show.legend = F) +
  scale_x_continuous(label = percent_format()) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Standard Deviation",
       y = "Expected Return",
       title = "Expected Monthly Return VS. Stand Deviation")

```

Let's direct visualized portfolio return and market return

```{r}
portfolio_mk_re %>%
  ggplot(aes(Rm, Ra)) +
  geom_point(color = "cornflowerblue") +
  labs(x = "Market Return",
       y = "Portfolio Return",
       title = "Scatterplot",
       caption = "Monthly balance portfolio, return monthly data") +
  geom_abline(aes(intercept = portfolio_beta$estimate[1],
                  slope = portfolio_beta$estimate[2]),
              color = "purple",
              size = .5) +
   geom_smooth(method = "lm", se = F, color = "green", size = .5) +
  scale_x_continuous(label = percent_format()) +
  scale_y_continuous(labels = percent_format())



```



```{r}
portfolio_augment <- portfolio_mk_re %>%
  do(model = lm(Ra ~ Rm, data = .)) %>%
  augment(model) %>%
  mutate(date = portfolio_re$date)

portfolio_augment

portfolio_augment %>%
  select(date, .fitted, Ra) %>%
  gather(.fitted, Ra, key = "type", value = "return") %>%
  ggplot(aes(date, return, color = type, group = type)) +
  geom_line() +
  scale_y_continuous(label = percent_format()) +
  labs(x = "", y = "Return", title = "Fitted VS actual Returns") +
  a_plex_theme() +
  a_main_fill() +
  a_main_color() +
  theme(legend.position = "bottom") 


```


```{r}

portfolio_augment %>%
  select(date, .fitted, Ra) %>%
  mutate(actual = 1000 * cumprod(1 + Ra),
         fitted = 1000 *cumprod(1 + .fitted)) %>%
  gather(actual, fitted, key = "type", value = "return") %>%
  ggplot(aes(date, return, group = type, color = type)) +
  geom_line() +
  labs(x = "", y = "", title = "Growth of a dollar: Actual VS Fitted",
       subtitle = "Initial Investment $1000") +
  scale_y_continuous(label = dollar_format()) +
  theme_tq()
  

```


```{r}
highchart() %>%
  hc_title(text = "Portfolio v. market returns") %>%
  hc_add_series(round(portfolio_mk_re$Ra, 4),
                        date = portfolio_mk_re$date,
                        type = "scatter") %>%
  hc_xAxis(title = list(text = "Market Returns")) %>%
  hc_yAxis(title = list(text = "Portfolio Returns")) %>%
  hc_tooltip(formatter = JS("function(){
                            return ('port return: ' + this.y + ' <br> mkt return: ' + this.x +  
                            ' <br> date: ' + this.point.date)}"))



```

## Fama Frech

FF multipel factor model of equity risk/return

[Fama French Data](http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html#HistBenchmarks)




```{r}

portfolio_re

file <- "/Users/yifeiliu/Documents/R/data/EDA/Fama_French/Global_3_Factors.csv"

three_factor <- read_csv(file, skip = 3)

three_factor <- three_factor %>%
  rename(date = X1,
         Mkt_rf = `Mkt-RF`) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(date =  rollback(ymd(parse_date_time(date, "%Y%m") + months(1))))



map_chr(three_factor, class)

```


```{r}
portfolio_ff <- portfolio_re %>%
  inner_join(three_factor, by = "date") %>%
  mutate(Mkt_rf = Mkt_rf / 100,
         SMB  = SMB / 100,
         HML = HML / 100,
         RF = RF / 100,
         r_excess = Ra - RF)

portfolio_ff
```


```{r}
ff_model <- portfolio_ff %>%
  do(model = lm(r_excess ~ Mkt_rf + SMB + HML, data = .)) %>%
  tidy(model, conf.int = T, conf.level = .95)

ff_model %>%
  mutate_if(is.numeric, ~ round(.,3)) %>%
  select(-statistic)

ff_model %>%
  mutate_if(is.numeric, funs(round(., 3))) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(term, y = estimate, shape = term, color = term)) +
  geom_point(show.legend = F) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), show.legend = F) +
  labs(title = "FF 3-Factor coefficient for our portfolio",
       x = "",
       y = "coefficient",
       caption = "data source: Fama French Website and Yahoo! Finance")

```


## Rolling Fama French

### get rolling lm
```{r}
window = 12

rolling_lm <- rollify(.f = function(r_excess, Mkt_rf, SMB, HML) {
  lm(r_excess ~ Mkt_rf + SMB + HML)
  },
  window = window, 
  unlist = F
)

```


### rolling betas

```{r}
rolling_ff_beta <- portfolio_ff %>%
  mutate(rolling_ff = rolling_lm(r_excess, Mkt_rf, SMB, HML)) %>%
  mutate(tidied = map(rolling_ff, tidy, conf.int = T)) %>%
  unnest(tidied) %>%
  slice(-1:-(window -1)) %>%
  rename(beta = estimate, factor = term) %>%
  select(date, factor, beta, conf.low, conf.high) %>%
  filter(factor != "(Intercept)") %>%
  group_by(factor)


rolling_ff_rsquared <- portfolio_ff %>%
  mutate(rolling_ff = rolling_lm(r_excess, Mkt_rf, SMB, HML)) %>%
  slice(-1:-(window -1)) %>%
  mutate(glanced = map(rolling_ff, glance)) %>%
  unnest(glanced) %>%
  select(date,r.squared, adj.r.squared, p.value) %>%
  mutate_if(is.numeric, ~ round(., digits = 4))

```

### Visulaizing Rolling Fama French

```{r}
# visualized rolling beta
rolling_ff_beta %>%
  ggplot(aes(date, beta, color = factor)) +
  geom_line() +
  labs(title= "12-Month Rolling FF Factor Betas",
       x = "rolling betas") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  theme_tq()


```

### Highchart
visualizing rolling r squared

```{r}
highchart(type = "stock") %>%
  hc_add_series(rolling_ff_rsquared, 
                hcaes(x = date, y = adj.r.squared),
                type = "line",
                color = "cornflowerblue",
                id = "R Sqauared") %>%
  hc_title(text = "Rolling Fmam French 3 Factor Adjusted R Square") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_yAxis(max = 1, min = 0) 
  # hc_add_theme(hc_theme_economist())
```
## Highcharting jobs friday

```{r}
codes_names_tbl <- tribble(
        ~ symbol, ~ names,
        "NPPTTL", "ADP Estimate",
        "PAYEMS", "Nonfarm Employment",
        "USCONS", "Construction",
        "USTRADE",   "Retail/Trade",
        "USPBS",  "Prof/Bus Serv",
        "MANEMP",    "Manufact",
        "USFIRE",    "Financial",
        "USMINE",   "Mining",
        "USEHS",    "Health Care",
        "USWTRADE",    "Wholesale Trade",
        "USTPU",    "Transportation",
        "USINFO",    "Info Sys",
        "USLAH",    "Leisure",
        "USGOVT",    "Gov",
        "USSERV",    "Other Services"
)

fred_empl_data <- tq_get(codes_names_tbl$symbol,                         
                          get = "economic.data",             
                          from = "2007-01-01")

fred_empl_data <- fred_empl_data %>%
  inner_join(codes_names_tbl, by = "symbol") %>%
  select(names, date, price) %>%
  rename(sector = names,
         employment = price)

```

show employment change over time

```{r}
emply_chagne <- fred_empl_data %>%
  group_by(sector) %>%
  mutate(monthly_change = employment - lag(employment)) %>%
  na.omit() %>%
  mutate(sign = case_when(
           monthly_change > 0 ~ "Positive",
           T ~ "Negative"))

emply_chagne %>%
  filter(sector == "Nonfarm Employment") %>%
  ggplot(aes(date, monthly_change, fill = sign)) +
  geom_col(alpha = 0.85) +
  scale_fill_manual(values = c("pink", "lightgreen")) +
  labs(title = "Monthly Private Employment Change",
       subtitle = "Total emp, since 2008",
       y = "Monthly chagne (thousands)",
       x = "",
       caption = "Data source: FRED | Author, @yifeiliu",
       fill = "") +
  theme_tq() 
  
emply_chagne %>%
  filter(sector == "Nonfarm Employment") %>%
  mutate(color_bar = case_when(sign == "Negative" ~ "#ffe6ea",
                               T ~ "#6495ed")) %>%
  hchart( type = "column", 
                pointWidth = 5,
                hcaes(x = date,
                      y = monthly_change,
                      color = color_bar),
                name = "monthly change")

```





## Resampling Fama French 

```{r}
ff_five <- read_csv("/Users/yifeiliu/Documents/R/data/EDA/Fama_French/Global_5_Factors_Daily.CSV",
                    skip = 3)

five_factor <- ff_five %>%
  rename(date = X1,
         Mkt_rf = `Mkt-RF`) %>%
  mutate(date = ymd(parse_date(as.character(date), "%Y%m%d"))) %>%
  mutate_if(is.numeric, funs(. / 100)) %>%
  select(-RF)


five_factor

stock_re_daily <- stock_price %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               type = "log",
               col_rename = "Ra") %>%
  group_by(symbol)  %>%
  filter(date >= "2013-01-01")

portfolio_five <- stock_re_daily %>%
  inner_join(five_factor, by = "date")
  
```





refernce linke: https://rviews.rstudio.com/categories/reproducible-finance-with-r/



```{r}
file_path <- "/Users/yifeiliu/Documents/R/data/EDA/sp500.xlsx"

cells <- xlsx_cells(file_path, sheets = "Ratios")
formats <- xlsx_formats(file_path)

select(cells, row, col, data_type, character, numeric, local_format_id)

bold <- formats$local$font$bold
indent <- formats$local$alignment$indent


cells %>%
  filter(row >= 12L,
         !is_blank) %>%
  behead("")
  behead("N", "time") %>%
  select(row, col, numeric, character, time)

```


