---
title: "IMDB analysis"
author: "Yifei Liu"
date: "11/20/2018"
output: html_document
---


```{r, message = F, warning=F}
library(R.utils) # unzip file
library(tidyverse) # tidy data analysis
library(tidytext)
library(naniar) # visualize missing data
library(ggplot2) # visualize data
library(scales) # change scales in x, y axis
library(lubridate)
library(Hmisc)
library(SDMTools)

detach("package:dplyr", unload=TRUE)
library(dplyr)

theme_set(theme_minimal()) # change ggplot theme to minimal
session_info()
```


Download IMDB dataset

```{r, eval=F}
file_dis <- "/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/IMDB_analysis"
setwd("/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/IMDB_analysis")
urls <- c("https://datasets.imdbws.com/name.basics.tsv.gz",
          "https://datasets.imdbws.com/title.akas.tsv.gz",
          "https://datasets.imdbws.com/title.basics.tsv.gz",
          "https://datasets.imdbws.com/title.crew.tsv.gz",
          "https://datasets.imdbws.com/title.episode.tsv.gz",
          "https://datasets.imdbws.com/title.principals.tsv.gz",
          "https://datasets.imdbws.com/title.ratings.tsv.gz")
dest_file <- c("name.basics.tsv.gz",
               "title.akas.tsv.gz",
               "title.basics.tsv.gz",
               "title.crew.tsv.gz",
               "title.episode.tsv.gz",
               "title.principals.tsv.gz",
               "title.ratings.tsv.gz")

for(i in seq_along(urls)) {
  download.file(urls[i], dest_file[i])
}

for (i in seq_along(dest_file)) {
  gunzip(dest_file[i])
}
setwd("/Users/yifeiliu/Documents/R/EDA")
```


Now we can read the file, we can load need data in R

```{r}
file_dis <- "/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/IMDB_analysis"
read_imdb <- function(data_path) {
  path <- "/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/IMDB_analysis/"
  read_tsv(paste0(path, data_path), na = "\\N", progress = F)
}

df_rating <- read_imdb("title.ratings.tsv")
df_principals <- read_imdb("title.principals.tsv")
df_basics <- read_imdb("title.basics.tsv")
df_actors <- read_imdb("name.basics.tsv")
df_episode <- read_imdb("title.episode.tsv")
df_akas <- read_imdb("title.akas.tsv")

```
## Popular TV Series Analysis

```{r}
df_basics %>%
  filter(titleType == "tvSeries",
         startYear <= 2018) %>%
  group_by(startYear) %>%
  summarize(observation = n()) %>%
  ggplot(aes(startYear, observation)) +
  geom_col()

```


We try to analyze the tv season data, so we need to aggreate the df_eposide data. 


```{r}
eposide_season <- df_episode %>%
  inner_join(df_rating) %>%
  group_by(parentTconst, seasonNumber) %>%
  summarize(averageRating = sum(averageRating * numVotes) / sum(numVotes),
            numVotes = sum(numVotes),
            eposide_num = max(episodeNumber)) %>%
  filter(!is.na(seasonNumber))


tv_eposides <- df_basics %>%
  right_join(eposide_season %>% rename(tconst = "parentTconst")) %>%
  select(-titleType, -endYear) %>%
  filter(startYear >= 1990,
         numVotes > 100) %>%
  group_by(tconst) %>%
  filter(1 %in% seasonNumber,
         all(seasonNumber == dplyr::row_number())) %>%
  ungroup()

us_tv <- df_akas %>%
  filter(region == "US") %>%
  inner_join(df_episode, by = c(titleId = "tconst")) %>%
  select(title, parentTconst)

tv_rating <- tv_eposides %>%
  filter(tconst %in% us_tv$parentTconst)



# new method

tv_basic <- df_episode %>%
  inner_join(df_basics) %>%
  select(- titleType, - endYear) %>%
  inner_join(df_rating, by = "tconst")


us_tv <- df_akas %>%
  rename(tconst = titleId) %>%
  filter(region == "US") %>%
  right_join(tv_basic, by = "tconst") %>%
  select(- language, - attributes, - isOriginalTitle)


# tv by season dataset

tv_season <- us_tv %>%
  group_by(seasonNumber, parentTconst, startYear, primaryTitle) %>%
  summarize(averageRating = sum(averageRating * numVotes) / sum(numVotes),
            numVotes = sum(numVotes, na.rm = T),
            runtime = mean(runtimeMinutes, na.rm = T),
            eposide_num = max(episodeNumber))


tv_season <- tv_season %>%
  group_by(parentTconst) %>%
  filter(startYear >= 1990,
         1 %in% seasonNumber,
         numVotes > 100) %>%
  mutate(year = startYear + seasonNumber - 1)

tv_season %>% filter(primaryTitle == "The Wire")


# join eposide with rating

test_1 <- df_episode %>%
  inner_join(df_rating)


# join test1 with us tv df_akas
test2 <- test_1 %>%
  inner_join(df_basics) %>%
  select(- titleType, - endYear)

# filter out non us tv series

us_tv <- df_akas %>%
  filter(region == "US") %>%
  select(titleId, title) %>%
  rename(tconst = titleId)

us_tv_eposide <- test2 %>%
  inner_join(us_tv, by = "tconst")

tv_season <- us_tv_eposide %>%
  group_by(parentTconst, seasonNumber, startYear) %>%
  summarize(averageRating = sum(averageRating * numVotes) / sum(numVotes),
            numVotes = sum(numVotes, na.rm = T),
            runtime = mean(runtimeMinutes, na.rm = T),
            eposide_num = max(episodeNumber))


```

Now we have complete tv show series start from 1985, only in US, number of total voters over 1000.



```{r}
tv_eposides %>%
  group_by(tconst) %>%
  filter(seasonNumber <= 5) %>%
  group_by(seasonNumber) %>%
  summarize(av_rating = mean(averageRating),
            num_voters = sum(numVotes))


tv_eposides %>%
  arrange(desc(averageRating))
```

what are the highest rating seasons of TV?

```{r}

by_show <- tv_eposides %>%
  group_by(tconst) %>%
  summarize(avg_rating = sum(averageRating * numVotes) / sum(numVotes),
            sd = wt.sd(averageRating, numVotes),
            seasons = n(),
            num_votes = sum(numVotes)) %>%
  arrange((avg_rating)) %>%
  inner_join(tv_eposides %>% select(tconst, primaryTitle)) %>%
  unique()

by_show %>%
  filter(seasons >= 5) %>%
  arrange(desc(sd))

tv_eposides %>%
  filter(primaryTitle %in% c("Scrubs", "Dexter")) %>%
  ggplot(aes(seasonNumber, averageRating, color = primaryTitle)) +
  geom_line()




```























Let's take a basic look at all those dataset and see what we can analysis


```{r}
gg_miss_var(df_rating, show_pct = T)
gg_miss_var(df_principals, show_pct = T)
gg_miss_var(df_basics, show_pct = T)
gg_miss_var(df_actors, show_pct = T)
```

we can first take a look at how many genre of film have been film in these years. 

```{r}
head(df_basics) 
count(df_basics, "titleType")

df_basics_processed <- df_basics %>%
  dplyr::filter(titleType == "movie",
                runtimeMinutes < 180,
                runtimeMinutes > 30,
                startYear < 2020) %>% 
  select(tconst, titleType, primaryTitle, startYear, runtimeMinutes, genres)

gg_miss_var(df_basics_processed, show_pct = T)

df_basics_processed %>%
  unnest_tokens(genre, genres, token = stringr::str_split, pattern = ",") %>%
  count("genre") %>%
  ggplot(aes(x = freq, y = reorder(genre, freq))) +
  geom_point() +
  labs(x = "",
       y = "Genre",
       title = "How many genre film been release since 1894",
       subtitle = "Data acquired from IMDB, Nov 20th, 2018")

```


## Mapping lead actors in movies

```{r}
df_actors_processed <- df_actors %>%
  dplyr::filter(str_detect(primaryProfession, "actor|actress")) %>%
  select(nconst, primaryName, birthYear, primaryProfession)


df_rating_basic <- df_rating %>%
  inner_join(df_basics_processed)

df_principals_processed <- df_principals %>%
  dplyr::filter(str_detect(category, "actor|actress")) %>%
  select(tconst, ordering, nconst, category) %>%
  group_by(tconst) %>%
  dplyr::filter(ordering == min(ordering))

df_movies <- df_principals_processed %>%
  inner_join(df_actors_processed) %>%
  inner_join(df_rating_basic)

dim(df_movies)
gg_miss_var(df_movies)

```




Reference:



