---
title: "NYT_style urban island maps"
author: "Yifei Liu"
date: "9/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages

```{r}
library(raster)
library(sf)
library(tidyverse)
library(ggtext)
library(showtext)
library(tigris)
font_add_google("Lato", regular.wt = 300, bold.wt = 700)

```


load data

```{r}
landsat_dc_july16 <- raster("/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/LST_F_20160702/LST_F_20160702.tif")

water <- st_read("https://opendata.arcgis.com/datasets/db65ff0038ed4270acb1435d931201cf_24.geojson") %>%
  st_transform(st_crs(landsat_dc_july16)) # use the same coordinate reference system as the landsat data

landsat_dc_july16

summary(landsat_dc_july16)

```

From this we can see that the temperature range from 70.5 to 100 with a median of 83 F and there are no NA

now covert the raster values to dataframe

```{r}
temp_spdf <- as(landsat_dc_july16, "SpatialPointsDataFrame")
temp_df <- as_tibble(temp_spdf)
colnames(temp_df) <- c("value", "x", "y")


water_sp <- as(water, "Spatial")

```


At this point we can make a "minmum viable product" map

```{r}
ggplot() +
  geom_raster(data = temp_df, aes(x = x, y = y, fill= value), interpolate = T) +
  geom_polygon(data = water_sp, aes(x = long, y = lat, group = group), color = "gray90", fill = "white") +
  coord_equal() +# make sure that the x and y axis are comparable
  theme_minimal()
```

Add some color scheme to my viz and set the breaks on the legend to the minimum, median, and maxium

```{r}

ggplot() +
  geom_raster(data = temp_df, aes(x, y = y, fill = value), interpolate = T) +
  geom_polygon(data = water_sp, aes(long, lat, group = group), color = "grey70", fill = "white") +
  theme_minimal() +
  coord_equal() +
  scale_fill_gradientn(colors = rcartocolor::carto_pal(name = "TealRose", n = 7), breaks = c(70, 83, 100), labels = c("70", "83", "100"), name = "")





```


Now we can reall start to see the differentiation across the city. The next step is to add the text neighborhood labels and descriptions in the margin. to do this I used the ggtext package which allows for markdown formatting, so it was easy to have bold characters. I created tibble with x/y coordiantes, and labels. To figure out the coordinates I used the axis as a strarting point and then trail and error from there. An easy way to get neighborhood label correct could have been use a neighborhood shapefile and use centroid from that, but you still need to figure out the description locations. 

```{r}
labels <- tibble(x = c(396000, 395900, 390500,
                       399700, 396000, 399700, 399200, 389070, 390000),
                 y = c(138300, 136600, 140000,
                       133000, 140000, 140500, 146400, 147000, 130100),
                 lab = c("W A S H I N G T O N", "downtown", "palisades", "anacostia", "columbia<br> heights", "brookland", "**Hotter:** Hotspots in<br>Brookland, Columbia<br>Heights, and LeDroit<br>Park hit **90 to 100 F**", "**Cooler:** Forested Rock<br>Creek Park recorded the<br>city's lowest temperatures<br>and helped to cool down<br>surrounding areas", "**Urban green spaces**<br>are an invaluable<br>resources for cooling<br>urban neighborhoods.<br>They help promote<br>walkability and improve<br>quality of living. Even a<br>few trees help!"))

ggplot() +
  geom_raster(data = temp_df, aes(x = x, y = y, fill = value, interpolate = T)) +
  geom_polygon(data = water_sp, aes(x = long, y = lat, group = group), color = "gray90", fill = "white") +
  geom_richtext(data = labels, aes(x = x, y = y, label = lab), fill = NA, label.color = NA, # remove backgroup and outline,
                label.padding = grid::unit(rep(0, 4), "pt"), color = "#656c78", hjust = 0) +
  theme_minimal() +
  coord_equal() +
  scale_fill_gradientn(colors = rcartocolor::carto_pal(name = "TealRose", n = 7), breaks = c(70, 83, 100), labels = c("70", "83", "100"), name = "")


```

Now the only thing left to do is clearing up the plot a bit and reformatting the legend. 

```{r}
ggplot() +
  geom_raster(data = temp_df, aes(x = x, y = y, fill = value, interpolate = T)) +
  geom_polygon(data = water_sp, aes(x = long, y = lat, group = group), color = "gray90", fill = "white") +
  geom_richtext(data = labels, aes(x = x, y = y, label = lab), fill = NA, label.color = NA, # remove backgroup and outline,
                label.padding = grid::unit(rep(0, 4), "pt"), color = "#656c78", hjust = 0) +
  theme_void() +
  coord_equal() +
  scale_fill_gradientn(colors = rcartocolor::carto_pal(name = "TealRose", n = 7), breaks = c(70, 83, 100), labels = c("70", "83", "100"), name = "") +
  labs(caption = "July 2, 2016\nSource: DC Open Data\nGraphic: Yifei Liu") +
  guides(fill = guide_colourbar(barheight = 0.3, barwidth = 20, direction = "horizontal", ticks = F)) +
  theme(legend.position = "bottom",
        text = element_text(color = "#656c78", family = "Lato"),
        plot.caption = element_text(hjust = 0))



```



## Try to reproduce this chart myself



```{r}
temp_sf <- spex::qm_rasterToPolygons(landsat_dc_july16)

ggplot() +
  geom_sf(data = temp_sf, aes(fill = LST_F_20160702), color = NA) +
  geom_sf(data = water,  color = "gray90", fill = "white") +
  theme_void() +
  coord_sf() +
  scale_fill_gradientn(colors = rcartocolor::carto_pal(name = "TealRose", n = 7), breaks = c(70, 83, 100), labels = c("70", "83", "100"), name = "")


```














[NYT-style urban heat island maps](https://www.katiejolly.io/blog/2019-08-28/nyt-urban-heat)











