---
title: "PG&E Energy Shutdown project"
author: "Yifei Liu"
date: "10/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(tidycensus)
library(rcartocolor)
library(leaflet)
library(sf)
library(zipcode)
library(leaflet.extras)
detach("package:dplyr", unload = TRUE)
library(dplyr)

```



Read files and process them and write new ones for flexdashboard ot use

```{r}

# PG&E shape file: https://www.pge.com/en_US/safety/emergency-preparedness/natural-disaster/wildfires/psps-event-maps.page


# cali_non <- read_sf("/Users/yifeiliu/Documents/R/data/book_exercise/geocom/PUBLIC_PSPS_10-09-19_D-02_BaseLine_1008_1039.kmz")
# cali <- read_sf("/Users/yifeiliu/Documents/R/data/book_exercise/geocom/PUBLIC_PSPS_10-10-19_B-07_T_24hrs_1010_0739.kmz")

# California population density data

# pge close location
pge <- st_read("/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/pge/pge_simple.gpkg")

# 2010 zip pop 
zip_pop <- read.table("/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/pge/Gaz_zcta_national.txt", header = T, colClasses = c("character", rep("numeric", 7))) %>%
  as_data_frame() %>%
  select(zip = GEOID, pop = POP10, household = HU10)
  
# all zip with state info
data("zipcode")

zip_cali <- zipcode %>%
  filter(state =="CA") %>%
  left_join(zip_pop, by = "zip") %>%
  select(zip, pop, household)

```


California zip code shapefile.

```{r}
# download data from https://catalog.data.gov/dataset/tiger-line-shapefile-2015-2010-nation-u-s-2010-census-5-digit-zip-code-tabulation-area-zcta5-na

pop_zip <- read_sf("/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/pge/tl_2015_us_zcta510.shp")

cali_zip <- zipcode %>%
  filter(state == "CA")

zip_area <- pop_zip %>%
  right_join(cali_zip, by = c("ZCTA5CE10" = "zip"))


zip_area_proj <- st_transform(zip_area, crs = 4326) # reproject

zip_simply <- rmapshaper::ms_simplify(zip_area_proj, keep = 0.1,
                                          keep_shapes = TRUE) 

cali_zip <- zip_simply %>%
  dplyr::select(ZCTA5CE10, city, geometry, latitude, longitude) %>%
  mutate(area = st_area(geometry))


affect_area <- st_intersection(cali_zip, pge) %>%
  mutate(effect_area = st_area(geometry))

cali_zip <- cali_zip %>%
  left_join(st_set_geometry(affect_area, NULL), by = c("ZCTA5CE10", "city", "area", "latitude", "longitude")) %>%
  mutate(Area = as.numeric(area) / 4046.86,
         Affect_area = as.numeric(effect_area) / 4046.86, 
         pct = Affect_area / Area * 100) %>%
  select(-area, -effect_area)
  

```


## web scrapping CA zip pop data

This method is prohibit by the dataset distributer, read the user aggrement. Keep in mind


## join zip pop and zip shapefile

```{r}
cali_comp <- cali_zip %>%
  left_join(zip_cali, by = c("ZCTA5CE10" = "zip"))

cali_df <- st_set_geometry(cali_comp, NULL) %>%
  mutate(affect = ifelse(is.na(Affect_area), "No", "Yes"),
         Affect_area = ifelse(is.na(Affect_area), "None Affected", round(Affect_area, 2)),
         pct = ifelse(is.na(pct), "None Affect", round(pct, 2)))

# write_sf(cali_comp, "/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/pge/cali_comp.gpkg")
# write_csv(cali_df, "/Users/yifeiliu/Documents/R/data/EDA/Blog_draft/pge/cali_comp.csv")



```

make a "ugly" static chart

```{r}
ggplot() +
  geom_sf(data = cali_comp, aes(fill = pop / 1000), color = "grey90") +
  geom_sf(data = pge, fill = "black", color = NA) +
  scale_fill_carto_c(palette = "PinkYl") +
  theme_void() +
  labs(fill = "Population(K)",
       title = "PG&E shut off power in California",
       subtitle = "Black area: Area have been turned off power by PG&E due to safety reason | Color palette: indicate population density",
       caption = "Data Source: PG&E | Graphic: Yifei Liu") +
  guides(fill = guide_colourbar(barheight = 0.3, barwidth = 20, direction = "horizontal", ticks = F)) +
  theme(legend.position = "bottom",
        text = element_text(family = "Lato", color = "#656c78"),
        plot.title = element_text(face = "bold", hjust = 0.5, size = 26),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, size = 12),
        plot.caption = element_text(hjust = 0))
```


write small files, easy for flexdashboard to process

## Interactive-ish map

```{r}
pal <- colorNumeric(palette = rcartocolor::carto_pal(7, "PinkYl") , NULL)

labs <- lapply(seq(nrow(cali_df)), function(i) {
  paste0('<p>', "City: ",cali_df[i, "city"],'<p></p>',
         "Zip Code: ",cali_df[i, "ZCTA5CE10"],'<p></p>',
         "Population: ",cali_df[i, "pop"], '<p></p>',
         "Affect or not: ", cali_df[i, "affect"], '<p></p>',
         "Affect Area: ", cali_df[i, "Affect_area"], " Acre",'<p></p>',
         "% of Area Affected: ", cali_df[i, "pct"])
})

leaflet(options = leafletOptions(dragging = F,
                                 minZoom = 5)) %>%
   addPolygons(data = pge, fillColor = "black", fillOpacity = 0.65, color = "black", 
              weight = 1) %>%
  addPolygons(data = cali_comp, stroke = FALSE, smoothFactor = 0.3, fillOpacity = 0.7, color = "#03F", 
              fillColor = ~pal(pop),
              label = lapply(labs, htmltools::HTML)) %>%
  addProviderTiles("OpenStreetMap.BlackAndWhite") %>%
  addResetMapButton()
```



## conclusion

- Name file is really really really hard........
- Google is God
- Best way to learn is read book + real word practice











