---
title: "student/teacher ratio"
author: "Yifei Liu"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(countrycode)
library(wbstats)
library(skimr)
library(WDI)
library(scales)
library(ggthemes)
library(snakecase)
theme_set(theme_minimal())

```

get the data,


I think we can got some other useful data for analysis. such as GDP PPP per capital from World bank. 

```{r}
student_ratio <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-07/student_teacher_ratio.csv")

WDIsearch("public.*education") %>%
  as.data.frame() %>%
  tbl_df() %>%
  arrange(str_length(name)) 

indicators_raw <- WDI(indicator = c("NY.GDP.PCAP.PP.KD","SE.ADT.LITR.ZS", "SP.POP.TOTL","SE.PRM.NENR.FE", "SE.PRM.NENR.MA"),
                      start = 2015, end = 2015, extra = T) %>%
  tbl_df()

indicators <- indicators_raw %>%
  select(country_code = iso3c,
         region,
         NY.GDP.PCAP.PP.KD:SE.PRM.NENR.MA) %>%
  mutate(country_code = as.character(country_code))


```

## EDA

```{r}
student_teacher_ratio_2015 <- student_ratio %>%
  filter(indicator == "Primary Education",
         year == 2015,
         !is.na(student_ratio))

student_teacher_ratio_2015 %>%
  arrange(desc(student_ratio)) %>%
  slice(1:10, seq(n() - 10, n())) %>%
  mutate(country = fct_reorder(country, student_ratio)) %>%
  ggplot(aes(country, student_ratio)) +
  geom_point() +
  coord_flip() +
  expand_limits(y = 0) +
  labs(title = "Countries with the highest and lowest student/teacher ratio")

```
we can see use the anti_join, we can observe that there are some data that exist in our _student_ratio_ but not in world bank PPP data, and some of these data just simply statistic summary of different region indicator such as world latin, asin etc. and others are just very small regions or country, we can just ignore it. Now we can join these two dataset, see the rationship between income and student/teacher ration.


```{r}

student_teacher_ratio_2015 %>%
  inner_join(indicators , by = "country_code") %>% 
  filter(!is.na(SE.ADT.LITR.ZS)) %>%
  mutate(literacy = SE.ADT.LITR.ZS / 100) %>%
  ggplot(aes(student_ratio, literacy)) +
  geom_point() +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  geom_text(aes(label = country), vjust = 1, just = 1, check_overlap = T)



```


```{r}
secondary_primary_education <- student_ratio %>%
  filter(year == 2015,
         !is.na(student_ratio),
         indicator %in% c("Primary Education", "Secondary Education")) %>%
  group_by(country) %>%
  # filter country that have both primary and secondary ed data
  filter(n() == 2) %>%
  ungroup()

secondary_primary_education %>%
  inner_join(indicators , by = "country_code") %>%
  filter(!is.na(NY.GDP.PCAP.PP.KD)) %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.PP.KD, 1/student_ratio)) +
  geom_point(aes(size = SP.POP.TOTL, fill = region), shape = 21, show.legend = F) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_size_continuous(range = c(0.5, 20)) +
  geom_text(aes(label = country), vjust = 1, just = 1, check_overlap = T) +
  labs(x = "Income level",
       y = "Teach to student ratio",
       title = "GDP per cap and teacher/student ratio are positive correlated",
       subtitle = "Income: GDP Per Cap, PPP 2010, Year = 2015") +
  theme(plot.title = element_text(face = "bold"))


```

### Other indicators

```{r}
joined <- student_teacher_ratio_2015 %>%
  inner_join(indicators, by = "country_code") %>%
  mutate(secondary_enrollment = (SE.PRM.NENR.FE + SE.PRM.NENR.MA) / 2)

joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(NY.GDP.PCAP.PP.KD, secondary_enrollment)) +
  geom_point() +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  labs(x = "GDP per capita",
       y = "secondary school enrollment") +
  geom_text(aes(label = country), check_overlap = T, hjust = 1, vjust = 1)

```

```{r}
joined %>%
  arrange(desc(SP.POP.TOTL)) %>%
  ggplot(aes(student_ratio, secondary_enrollment)) +
  geom_point() +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  labs(x = "Student/teacher ratio in primary school",
       y = "secondary school enrollment") +
  geom_text(aes(label = country), check_overlap = T, hjust = 1, vjust = 1)

```

Examining the confounding variables:

```{r}

joined %>%
  lm(secondary_enrollment ~ student_ratio + log2(NY.GDP.PCAP.PP.KD), 
     data = .) %>%
  summary()


```


```{r}
joined %>%
  transmute(student_ratio,
            secondary_enrollment,
            log2_gpd = log2(NY.GDP.PCAP.PP.KD)) %>%
  cor(use = "pairwise.complete.obs") %>%
  eigen()



```

we run into multicollinearity problem in this case. 











