---
title: "College Major"
author: "Yifei Liu"
date: "12/14/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plotly)
library(tidyverse)
library(scales)
library(ggrepel)
library(broom)
library(dplyr)

theme_set(theme_minimal())
```


```{r}
recent_grads <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-16/recent-grads.csv")

# we didn't get this dataset first, we get this data by plot the raw data frist, and figure out how we what processed the data
majors_processed <- recent_grads %>%
  arrange(desc(Median)) %>%
  mutate(Major = str_to_title(Major),
        Major = fct_reorder(Major, Median))
```

I'll also be aggregating by cateogyr. 


```{r}
by_majoy_cateotry <- majors_processed %>%
  dplyr::filter(!is.na(Total)) %>%
  group_by(Major_category) %>%
  dplyr::summarize(Men = sum(Men),
               Women = sum(Women),
               Total = sum(Total),
            MedianSalary = sum(Median * Sample_size) / sum(Sample_size)) %>%
  mutate(ShareWomen = Women / Total) %>%
  arrange(desc(ShareWomen))

```

### Most common Majors


What major categories (e.g. enginerring, psycholog, business) were most common?

```{r}
by_majoy_cateotry %>%
  mutate(Major_category = fct_reorder(Major_category, Total)) %>%
  gather(Gender, Number, Men, Women) %>%
  ggplot(aes(Major_category, Number, fill = Gender)) +
  geom_col() +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() +
  labs(title = "What are the most common major category",
       x = "",
       y = "Total # of graduess")


```

### what categories of majors make more money than others

```{r}
recent_grads %>%
  mutate(Major_category = fct_reorder(Major_category, Median)) %>%
  ggplot(aes(Major_category, Median, fill = Major_category)) +
  geom_boxplot() +
  scale_y_continuous(labels = dollar_format()) +
  expand_limits(y = 0) +
  coord_flip() +
  theme(legend.position = "none")
```





### what are the highest earning majors?

```{r}
majors_processed %>%
  head(20) %>%
  ggplot(aes(Major,Median, color = Major_category)) +
  scale_y_continuous(labels = dollar_format()) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  coord_flip() +
  expand_limits(y = 0)

majors_processed %>%
  filter(Sample_size >= 100) %>%
  head(20) %>%
  ggplot(aes(Major,Median, color = Major_category)) +
  scale_y_continuous(labels = dollar_format()) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  coord_flip() +
  expand_limits(y = 0) +
  labs(title = "What are the highest-earning majors",
       subtitle = "Top 20 majors with at least 100 graduates survey. Bars represent the 25th to 75th percentage",
       x = "",
       y = "median salary graduates")

```





### How does gender breakdown related to typical earnings?

```{r}
majors_processed %>%
  arrange(desc(Total)) %>%
  head(20) %>%
  gather(Gender, Number, Men, Women) %>%
  mutate(Major = fct_reorder(Major, Total)) %>%
  ggplot(aes(Major, Number, fill = Gender)) +
  geom_col() +
  coord_flip()
```





```{r}
by_majoy_cateotry %>%
  ggplot(aes(ShareWomen, MedianSalary)) +
  geom_point() +
  geom_text_repel(aes(label = Major_category), force = .2) +
  geom_smooth(method = "lm") +
  expand_limits(y = 0) +
  scale_y_continuous(labels = dollar_format())


```


When plot the correlation, you probabily don't want to aggregate first. If you do you will loose a lot of information about variation. Keep us safe from *Simpson’s Paradox*. What problem could rise if we aggreate data first. For example, we can see in last chart, the health is the pretty high in terms of median income and women percentage. What we also know nurse also pretty high in women percentage but pretty low in terms of income level. Correlation within each category that are overall different than overall correlation. 


```{r}
g <- majors_processed %>%
  filter(!is.na(Total)) %>%
  mutate(Major_category = fct_lump(Major_category, n = 4)) %>%
  ggplot(aes(ShareWomen, Median, color = Major_category, size = Sample_size, label = Major)) +
  geom_point() +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  geom_smooth(method = "lm", aes(group = 1)) +
  expand_limits(y = 0)

ggplotly(g)

```



```{r}
majors_processed %>%
  select(Major, Total, ShareWomen, Sample_size, Median) %>%
  lm(Median ~ ShareWomen, data =., weights = Sample_size) %>%
  summary()

```


```{r}
majors_processed %>%
  select(Major, Major_category, Total, ShareWomen, Sample_size, Median) %>%
  add_count(Major_category) %>%
  filter(n >= 10) %>%
  nest(-Major_category) %>%
  mutate(model = map(data, ~lm(Median ~ ShareWomen, data = ., weights = Sample_size)),
         tidied = map(model, tidy)) %>%
  unnest(tidied) %>%
  filter(term == "ShareWomen") %>%
  arrange(estimate) %>%
  mutate(fdr = p.adjust(p.value, method = "fdr"))
  
```

```{r}
majors_processed %>%
  filter(Sample_size >= 100) %>%
  mutate(IQR = P75th - P25th) %>%
  arrange(desc(IQR))
  

```


### Futuer work

* Examining Unemployment and fraction taking a job requrieing a college degree
* Examining interquartile ranges







### Appendix

```{r}
majors_processed %>%
  ggplot(aes(Sample_size, Median)) +
  geom_point() +
  geom_text(aes(label = Major), check_overlap = T, vjust = 1, hjust = 1) +
  scale_x_log10()
```




```{r}
knitr::knit_exit()
```


What were the msot common *majors*? (Since there wer 173, we are not going to show them all)

```{r}
majors_processed %>%
  mutate(Major = fct_reorder(Major, Total)) %>%
  arrange(desc(Total)) %>%
  head(20) %>%
  ggplot(aes(Major, Total, fill = Major_category)) +
  geom_col() +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() +
  labs(x = "",
       y = "Total # of graduess")


```



```{r}

recent_grads %>%
  group_by(Major_category) %>%
  summarize(Median = median(Median)) %>%
  mutate(Major_category = fct_reorder(Major_category, Median)) %>%
  ggplot(aes(Major_category, Median)) +
  geom_col() +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip()
  
```

### What are the lower earning Majors

```{r}
majors_processed %>%
  filter(Sample_size >= 100) %>%
  tail(20) %>%
  ggplot(aes(Major,Median, color = Major_category)) +
  scale_y_continuous(labels = dollar_format()) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  coord_flip() +
  expand_limits(y = 0)
```















