---
title: "Malaria EDA"
author: "Yifei Liu"
date: "11/13/2018"
output: html_document
---

```{r}
library(tidyverse)
library(malariaAtlas)
library(ggplot2)
library(gganimate)
library(fuzzyjoin)
library(dplyr)
theme_set(theme_light())
```

```{r}
listData(datatype = "pr points")
countries_pr <- tbl_df(getPR(country = "ALL", species = "both"))

# take a look which country have the most available data
countries_pr %>%
  group_by(country) %>%
  dplyr::summarise(n = n()) %>%
  arrange(desc(n))

# let's study indonesia
Indonesia_pr %>%
  dplyr::filter(is.na(pr)) %>%
  dim()
# we can see how many data was unavaibale due to permission issues

Indonesia_pr <- tbl_df(getPR(country = "Indonesia", species = "both"))
Indonesia_pr <- Indonesia_pr %>%
  dplyr::filter(!is.na(pr))
```




```{r}

Indonesia_pr %>%
  group_by(year_start) %>%
  dplyr::summarise(examined = sum(examined),
                   positive = sum(positive),
                   study = n()) %>%
  mutate(pr = positive / examined) %>%
  ggplot(aes(year_start, pr)) +
  geom_line()
Indonesia_pr %>%
  mutate(decade = 10 * (year_end %/% 10)) %>%
  arrange(pr) %>%
  ggplot(aes(longitude, latitude, colour = pr)) +
  borders("world", regions = "Indonesia") +
  geom_point() +
  scale_color_gradient2(low = "blue", high = "red", midpoint = .30, labels = scales::percent_format()) +
  facet_wrap(~decade) +
  theme_void() +
  coord_map() +
  labs(color = "Prevalence")

```

We can by plot these two chart side by side and find, even though in second chart, thoughtout the decades we can see a decline of PR rate reduction from 2000 to 2010, but when also consider chart 1, we can see overall avaiable research goes down. So we cannot conclude malaria goes down. 


## Aggregated across countries

```{r}
malaria_inc <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-11-13/malaria_inc.csv")

malaria_inc_processed <- malaria_inc %>%
  setNames(c("country", "code", "year", "incidence")) %>%
  mutate(incidence = incidence / 1000)


```

```{r}
malaria_inc_processed %>%
  dplyr::filter(country %in% sample(unique(country), 6)) %>%
  ggplot(aes(year, incidence, color = country)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format())
```
we may want see the extreme data such as during those years which country decrease the incidence the most. Look at 2015 leves and the change from 2015 to 2000
```{r}
malaria_spread <- malaria_inc_processed %>%
  mutate(year = paste0("Y", year)) %>%
  spread(year, incidence)




malaria_spread %>%  dplyr::filter(country != "Turkey",
                !is.na(code)) %>%
  mutate(current = Y2015,
         change = Y2015 - Y2000) %>%
  ggplot(aes(current, change)) +
  geom_point() +
  geom_text(aes(label = code), vjust = 1, hjust = 1)
```






```{r}
malaria_spread %>%
  anti_join(map_data("world"), by = c(country = "region"))


malaria_inc_processed %>%
  dplyr::filter(incidence < 1) %>%
  inner_join(maps::iso3166 %>%
               select(a3, mapname), by = c(code = "a3")) %>%
  inner_join(world, by = c(mapname = "region")) %>%
  ggplot(aes(long, lat, group = group, fill = incidence)) +
  geom_polygon() +
  coord_map() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = .2, labels = scales::percent_format()) +
  facet_wrap(~ year) +
  theme_void() +
  labs(title = "Malaria Incidence over time around the world")

```


### Malaria Deaths over time

```{r}
malaria_deaths <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-11-13/malaria_deaths.csv")

malaria_deaths_processed <- malaria_deaths %>%
  setNames(c("country", "code", "year", "deaths"))

```


```{r}



world <- map_data("world") %>%
  dplyr::filter(region != "Antarctica")

malaria_deaths_processed %>%
  inner_join(maps::iso3166 %>%
               select(a3, mapname), by = c(code = "a3"))%>%
  mutate(mapname = str_remove(mapname, "\\(.*")) %>%
  mutate(continent = countrycode(code, "iso3c", "continent")) %>%
  dplyr::count(continent, wt = deaths)


malaria_country_data <- malaria_deaths_processed %>%
  inner_join(maps::iso3166 %>%
               select(a3, mapname), by = c(code = "a3"))%>%
  mutate(mapname = str_remove(mapname, "\\(.*"))

malaria_map_data <- world %>%
  inner_join(malaria_country_data, by = c(region = "mapname")) %>%
  tbl_df()

malaria_map_data %>%
  ggplot(aes(long, lat, group = group, fill = deaths)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 100) +
  theme_void() +
  labs(title = "Malaria Deaths over time around the world",
       fill = "Death Per 100,000")


```


```{r}

malaria_map_data %>%
  mutate(continent = countrycode(code, "iso3c", "continent")) %>%
  dplyr::filter(continent == "Africa") %>%
  dplyr::filter(year %% 2 == 0) %>%
  ggplot(aes(long, lat, group = group, fill = deaths)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 100) +
  theme_void() +
  labs(title = 'Malaria Deaths over time in Africa',
       fill = "Death Per 100,000") +
  transition_manual(year) +
  ease_aes('linear')




```






