---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F)

```



```{r}
library(tidyverse)
library(scales)
library(splines)
library(broom)
theme_set(theme_minimal())

maryland_bridges <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-11-27/baltimore_bridges.csv") %>%
  select(- vehicles) %>%
  mutate(inspection_yr = inspection_yr + 2000,
         decade = 10 * (yr_built %/% 10),
         responsibility = fct_lump(responsibility, 4),
         county = str_to_title(county))

```

```{r}

maryland_bridges %>%
  dplyr::filter(yr_built >= 1900) %>%
  dplyr::count(decade = 10 * (yr_built %/% 10), sort = T) %>%
  ggplot(aes(decade, n)) +
  geom_line() +
  expand_limits(y = 0) +
  labs(y = "# of bridges in Baltimore buit this decade")
    


```



How does the condition of brudges depend on how old it is?


```{r}

maryland_bridges %>% 
  dplyr:::filter(yr_built >= 1900) %>%
  group_by(decade) %>%
  dplyr::summarize(pct_good = mean(bridge_condition == "Good"),
                   total = n()) %>%
  ggplot(aes(decade, pct_good)) +
  geom_line() +
  scale_y_continuous(labels = percent_format())
  



```





Most bridges built before 1970 we could not cosider in "good" condition (thogh it doesn't matter how long before 1970 they were built) while a majority of bridges built since 2000 are in good condition. 


```{r}
maryland_bridges %>%
  replace_na(list(responsibility = "Other")) %>%
  dplyr::count(responsibility = fct_lump(responsibility, 4), sort = T) %>%
  mutate(responsibility = fct_reorder(responsibility, n)) %>%
  ggplot(aes(responsibility, n)) +
  geom_col() +
  coord_flip()
  
```



```{r}
maryland_bridges %>%
  dplyr:::filter(yr_built >= 1900) %>%
  replace_na(list(responsibility = "Other")) %>%
  group_by(responsibility = fct_lump(responsibility, 4),
               decade) %>%
  dplyr::summarize(pct_good = mean(bridge_condition == "Good"),
                   total = n()) %>%
  dplyr::filter(responsibility != "Other") %>%
  ggplot(aes(decade, pct_good, color = responsibility)) +
  geom_line() +
  scale_y_continuous(labels = percent_format()) +
  expand_limits(y = 0) +
  labs(y = "% of bridges rated `Good`")
```


The County Highway agency consisitenly has the most bridges rated as "Good" from before 1970. State toll Authority is underperforming in bridges built in the 1990s.
```{r}

maryland_bridges %>%
  ggplot(aes(avg_daily_traffic)) +
  geom_histogram() +
  scale_x_log10(labels = comma_format())

```


```{r}
maryland_bridges %>%
  dplyr::filter(yr_built >= 1990) %>%
  group_by(traffic_category = cut(avg_daily_traffic, c(0, 1000, 10000, Inf),
                                  labels = c("<1000", "1000- 10,000", "10,000 +"))) %>%
  dplyr::summarize(pct_good = mean(bridge_condition == "Good"),
                   total = n())
```


What does the traffic look like geographically?

```{r}
maryland_bridges %>%
  ggplot(aes(long, lat, color = avg_daily_traffic)) +
  borders("state", regions = "Maryland") +
  geom_point() +
  theme_void() +
  coord_map() +
  scale_color_gradient2(low = "blue", high = "red", midpoint = log10(median(maryland_bridges$avg_daily_traffic)), trans = "log10",
                        labels = comma_format())

```


```{r}
maryland_bridges %>%
  ggplot(aes(long, lat, color = bridge_condition)) +
  borders("state", regions = "Maryland") +
  geom_point(size = 1) +
  theme_void() +
  coord_map()
```




```{r}
maryland_bridges %>%
  ggplot(aes(long, lat, color = bridge_condition)) +
  borders("state", regions = "Maryland") +
  geom_point(size = 1) +
  theme_void() +
  coord_map() 
```


```{r}
maryland_bridges %>%
  dplyr::filter(yr_built >= 1900) %>%
  ggplot(aes(long, lat, color = yr_built)) +
  borders("state", regions = "Maryland") +
  geom_point(size = 1) +
  theme_void() +
  coord_map() 
```

```{r}
maryland_bridges %>%
  dplyr::filter(yr_built >= 1900) %>%
  ggplot(aes(long, lat, color = county)) +
  borders("state", regions = "Maryland") +
  geom_point(size = 1) +
  theme_void() +
  coord_map() 
```

```{r}
maryland_bridges %>%
  dplyr:::filter(yr_built >= 1900) %>%
  group_by(county,  decade) %>%
  dplyr::summarize(pct_good = mean(bridge_condition == "Good"),
                   total = n()) %>%
  arrange(county, decade) %>%
  View()
  ggplot(aes(decade, pct_good, color = county)) +
  geom_line() +
  scale_y_continuous(labels = percent_format()) +
  expand_limits(y = 0) +
  labs(y = "% of bridges rated `Good`")
```


### Effect of county, responsibility and time on bridge condition


```{r}

bridges <- maryland_bridges %>%
  dplyr::filter(yr_built >= 1900)

# fit a logistic regress

simple_model <- bridges %>%
  dplyr::mutate(good = bridge_condition == "Good") %>%
  glm(good ~ ns(yr_built, 4), data = ., family  = "binomial")


augment(simple_model, bridges, type.predit = "response") %>%
  ggplot(aes(yr_built, .fitted)) +
  geom_line() +
  scale_y_continuous(labels = percent_format()) +
  labs(y = "Predicted Probability a bridge is rated `Good`")


model <- bridges %>%
  dplyr::mutate(good = bridge_condition == "Good") %>%
  glm(good ~ ns(yr_built, 4) + responsibility + county, data = ., family  = "binomial")
  

augment(model, bridges, type.predit = "response") %>%
  ggplot(aes(yr_built, .fitted, color = responsibility)) +
  facet_wrap(~county) +
  geom_line() +
  scale_y_continuous(labels = percent_format()) +
  labs(y = "Predicted Probability a bridge is rated `Good`")

```

Controlled for the overall trend in change over time, what's the effect of being in a particular county or responsibility?


```{r}
model %>%
  tidy(conf.int = T) %>%
  dplyr::filter(str_detect(term, "responsibility|county")) %>%
  mutate(term = reorder(term, estimate)) %>%
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, color = "red", lty = 2)


```





We haven't found evidence of an effect of geography or ownership on bridge condition, once we control for time































