---
title: "Bob Ross Painting"
author: "Yifei Liu"
date: "8/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load the packages and dataset

```{r}
library(tidyverse)
library(widyr)
library(ggraph)
library(igraph)
library(reshape2)
library(tidytext)
library(broom)

custome_theme <- theme_minimal(base_family = "Space Mono") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(face = "italic"))

theme_set(custome_theme)

bob_ross <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-06/bob-ross.csv") %>%
  janitor::clean_names() %>%
  tidyr::extract(episode, into = c("season", "episode_number"), "S(.*)E(.*)", convert = T, remove = F)

```


```{r}
bob_ross_gathered <- bob_ross %>%
  gather(element, present, -season, -episode_number, -title, -episode) %>%
  filter(present == 1) %>%
  mutate(title = str_to_title(str_remove_all(title, '"')),
         element = str_to_title(str_replace(element, "_", " "))) %>%
  select(-present)

```


```{r}
bob_ross_gathered %>%
  count(element, sort = T) %>%
  head(25) %>%
  mutate(element = fct_reorder(element, n)) %>%
  ggplot(aes(element, n, fill = element)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "", y = "")
  

```


```{r}
bob_ross_gathered %>%
  add_count(episode) %>%
  arrange(desc(n)) %>%
  View()

```


How have Ross's paintings been changing over time

```{r}
by_season_element <- bob_ross_gathered %>%
  filter(!element %in% c("Tree", "Trees")) %>%
  group_by(season) %>%
  mutate(number_episodes = n_distinct(episode)) %>%
  count(season, element, number_episodes, sort = T) %>%
  mutate(percent_included = n / number_episodes) %>%
  group_by(element) %>%
  mutate(element_total = sum(n)) %>%
  ungroup()


by_season_element %>%
  filter(element_total >= 100) %>%
  ggplot(aes(season, percent_included, color = element)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format()) +
  expand_limits(y = 0) +
  facet_wrap(~element)


```


Could have used: [multiple models with broom](https://r4ds.had.co.nz/many-models.html)



### Clustering

What tends to appear together

```{r}
correlations <- bob_ross_gathered %>%
  add_count(element) %>%
  filter(n >= 10) %>%
  pairwise_cor(element, episode, sort = T)


correlations %>%
  filter(item1 == "River") %>%
  mutate(item2 = fct_reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation, fill = correlation > 0)) +
  geom_col(aes(fill = correlation > 0), show.legend = F) +
  coord_flip() +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  labs(y = "",
       x = "",
       title = "What tends to appear with a River",
       subtitle = "Among elements that appeared in at least 10 times")

```


```{r}
correlations %>%
  head(100) %>%
  graph_from_data_frame() %>%
  ggraph() +
  geom_edge_link(aes(alpha = correlation)) +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()



```

### Principle Component Analysis

what dimensions drive a lot of the variation among paintings?

```{r}

binary_matrix <- bob_ross_gathered %>%
  acast(title ~ element)

# center the columns

centered_matrix <- t(t(binary_matrix) - colMeans(binary_matrix))

svd_result <- svd(centered_matrix)

element_weights <- tidy(svd_result, matrix = "v") %>%
  mutate(element = colnames(binary_matrix)[column])

element_weights %>%
  filter(PC <= 4) %>% 
  group_by(PC) %>%
  top_n(20, abs(value)) %>%
  ungroup() %>%
  mutate(element = reorder_within(element, value, PC)) %>%
  ggplot(aes(element, value)) +
  geom_col(aes(fill = value > 0), show.legend = F) +
  coord_flip() +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  facet_wrap(~ PC, scales = "free") +
  labs(x = "", y = "",
       title = "First four principle components for Bob Ross Paintings") 
  

```

1. Mountain vs Ocean/Beach and deciduous trees
2. Trees, especially deciduious, vs Ocean
3. Spring/Summer vs Winter
4. Lake vs River


```{r}
painting_weights <- tidy(svd_result, matrix = "u") %>%
  mutate(painting = rownames(binary_matrix)[row])

```

we can ask:

what is the painting is the strongest in principle component 1

```{r}
painting_weights %>%
  filter(PC == 1) %>%
  arrange(desc(value))


bob_ross_gathered %>%
  filter(title == "Frozen Solitude")

painting_weights %>%
  filter(PC <= 4) %>% 
  group_by(PC) %>%
  top_n(20, abs(value)) %>%
  ungroup() %>%
  mutate(painting = reorder_within(painting, value, PC)) %>%
  ggplot(aes(painting, value)) +
  geom_col(aes(fill = value > 0), show.legend = F) +
  coord_flip() +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  facet_wrap(~ PC, scales = "free") +
  labs(x = "", y = "",
       title = "First four principle components for Bob Ross Paintings") 


```

```{r}
tidy(svd_result, matrix = "d") %>%
  ggplot(aes(PC, percent)) +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "", y = "")

```





















