---
title: 'TidyTuesday: Car fule efficiency'
author: "Yifei Liu"
date: "10/15/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load pacakges and files

```{r}
library(tidyverse)
library(hrbrthemes)
library(ggthemes)
library(ggeconodist)
library(broom)
library(splines)
library(ggcorrplot)
detach("package:dplyr", unload = TRUE)
library(dplyr)
library(gganimate)
library(ggeconodist)
library(showtext)
library(waffle)
theme_set(theme_minimal())
font_add_google("Lato", regular.wt = 300, bold.wt = 700)

```

```{r}
big_epa_cars <- read_csv("https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-10-15/big_epa_cars.csv?raw=true") %>%
  mutate(uses_electricity = ifelse(highwayE > 0, "Electric Vehicles", "Fossil Fuel Vehicles"))


```

# David Robinson Analysis

```{r}

# rearrange column by alphabetically
cars_alpha <- big_epa_cars %>%
  select(sort(colnames(big_epa_cars))) 

cars_alpha %>%
  count(atvType, sort = T)

```



```{r}

big_epa_cars %>%
  ggplot(aes(highway08, city08)) +
  geom_point(alpha = 0.3) +
  geom_abline(color = "red") +
  expand_limits(x = 0, y = 0) +
  facet_wrap(~ uses_electricity, scales = "free") +
  labs(x = "Higway MPG",
       y = "City MPG",
       title = "How does fuel efficiency differ between city + highway") +
  theme_fivethirtyeight()
  

```


```{r}

big_epa_cars %>% 
  select(city08, highway08, make, model, cylinders, displ, drive, engId, eng_dscr)

```

```{r}

big_epa_cars %>%
  filter(cityE == 0) %>%
  mutate(VClass = fct_lump(VClass, 8),
         VClass = fct_reorder(VClass, city08)) %>%
  ggplot(aes(VClass, city08)) +
  geom_boxplot() +
  coord_flip()

big_epa_cars %>%
  filter(cityE == 0) %>%
  mutate(drive = fct_reorder(drive, city08)) %>%
  ggplot(aes(drive, city08)) +
  geom_boxplot() +
  coord_flip()

big_epa_cars %>%
  filter(cityE == 0) %>%
  ggplot(aes(cylinders, city08, group = cylinders)) +
  geom_boxplot() 

big_epa_cars %>%
  filter(cityE == 0) %>%
  ggplot(aes(displ, city08, group = displ)) +
  geom_point()+
  expand_limits(x = 0, y = 0)



```


## Goal: predict city fuel efficiency for single-fuel, non-electric efficiency cars

```{r}

# Cross validation holdout se
non_electric_cars <- big_epa_cars %>%
  filter(cityA08 == 0, cityE == 0) %>%
  sample_frac(1)

training_set <- non_electric_cars %>%
  filter(row_number() %% 5 != 0)


```



```{r}
training_set %>%
  ggplot(aes(displ, city08)) +
  geom_point() +
  geom_smooth(method = "lm")

# simple linear regression not work very work very well, we can clear see trend in residuals

lm(city08 ~ displ, data = training_set) %>%
  augment(data = training_set) %>%
  ggplot(aes(displ, .resid)) +
  geom_point() 

# add Natural Cubic splines to fit linear regression model, see what happen.

augmented_data <- lm(city08 ~ ns(displ, 2), data = training_set) %>%
  augment(data = training_set) 

# what model look like

lm(city08 ~ ns(displ, 2), data = training_set) %>% tidy()

augmented_data %>%
  ggplot(aes(displ, city08)) +
  geom_point() +
  geom_line(aes(y = .fitted), color = "red", size = 2)


models <- tibble(df = 1:6) %>%
  mutate(lm_model = map(df, ~ lm(city08 ~ ns(displ, df = .), data = training_set)))

augmented_unnested <- models %>%
  mutate(augmented = map(lm_model, augment, data = training_set)) %>%
  unnest(augmented) 

augmented_unnested %>%
  ggplot(aes(displ, city08)) +
  geom_point(data = training_set) +
  geom_line(aes(y = .fitted, color = factor(df)), size = 2) +
  labs(x = "Engine volumn(L)",
       y = "City MPG",
       color = "# of degrees of freedom") +
  expand_limits(x = 0, y = 0) +
  theme(legend.position = "bottom")

augmented_unnested %>%
  ggplot(aes(displ, .resid)) +
  geom_point() +
  facet_wrap(~ df) # base on the residuals chart, we don't find the model improve that much. 

# contrary to augment which give you estimate parameter of models, glance give you models' goodness of fitness
glanced_models <- models  %>%
  rename(spline_df = df) %>%
  mutate(glanced = map(lm_model, glance, data = training_set)) %>%
  unnest(glanced)

glanced_models %>%
  ggplot(aes(spline_df, adj.r.squared)) +
  geom_line() # the chart show add degree of freedom increase r values, but plateau around 4

```



```{r}

lm(city08 ~ ns(displ, 4) + ns(cylinders, 3), data = training_set) %>%
  anova() %>%
  tidy() %>%
  mutate(pct_variation = sumsq / sum(sumsq)) # due to high correlation between cylinders and displ

```



```{r}

efficiency_time <- training_set %>%
  mutate(VClass = fct_lump(VClass, 6),
         guzzler = guzzler %in% c("G", "T")) %>%
  group_by(year = 2 * floor(year / 2), VClass) %>%
  summarize(avg_efficiency = mean(city08, na.rm = T),
            avg_cyl = mean(cylinders, na.rm = T)) %>%
  summarize_at(vars(city08, cylinders, displ, guzzler), mean)

efficiency_time %>%
  ggplot(aes(year, avg_efficiency, color = VClass)) +
  geom_line() +
  expand_limits(y = 0)  
# we can see midsize and large size cars efficiency increase the most

efficiency_time %>%
  ggplot(aes(year, avg_cyl, color = VClass)) +
  geom_line() +
  expand_limits(y = 0)
# we can see large size cars cylinders number drop
  
```


# My own analysis /Vis

## Car CO2 emission trend

see how co2 emission cross different size vehicles change over past two decades
```{r}

dim(big_epa_cars)

```

we have 84 varaibels and 41804 observations

```{r}
skimr::skim(big_epa_cars) 
```

most numeric variable are complete, over 50% of character variables are also complete. But most logical variables are missing.

```{r}
big_epa_cars %>%
  filter(uses_electricity == "Fossil Fuel Vehicles") %>%
  count(fuelType, fuelType1, fuelType2, sort = T) # we don't have fueltype2

big_epa_cars %>%
  filter(uses_electricity == "Fossil Fuel Vehicles") %>%
  ggplot() + 
  geom_point(aes(x = co2, y = co2TailpipeGpm))

```

we still have a lot of co2 data = -1 and 0 , = 0 case is the car use electricity. = -1, Need further inspection. We have about 9490 point  = -1, about 23% of entire non-ele dataset

```{r}

big_epa_cars %>%
  filter(uses_electricity == "Fossil Fuel Vehicles") %>%
  count(co2 < 0)

big_epa_cars %>%
  filter(uses_electricity == "Fossil Fuel Vehicles") %>%
  count(co2TailpipeGpm < 0)



```

now a lot of co2 values is less than zero but another co2 measuremnet co2TailpipeGpm is more complete so we use these variables for co2 benchmark


check emission levels

```{r}

car_emission <- big_epa_cars %>%
  filter(uses_electricity == "Fossil Fuel Vehicles")

car_emission %>%
  group_by(year) %>%
  summarize(co2_median = median(co2TailpipeGpm, na.rm = T)) %>%
  ggplot(aes(year, co2_median)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", color = "#589bb7", fill = "#589bb7")  +
  labs(x = "", y  = "CO2 Emission level (grams/miles)",
       title = "Average car CO2 emission",
       caption = "Source: EPA") +
  theme_ipsum_rc(grid = "Y")
  


```

see what type of cars emission change during last 20+ years

```{r}

name_clean <- c("\\ -.*", " Cars", " Vehicle", " Trucks")

car_emission <- car_emission %>%
  mutate(VClass = fct_lump(VClass, 6),
         VClass = str_remove_all(VClass, paste(name_clean, collapse = "|")),
         year = 2 * floor(year / 2)) %>%
  group_by(year, VClass) %>%
  summarize(co2_mean = mean(co2TailpipeGpm, na.rm = T))

p <- car_emission %>%
  ggplot(aes(year, co2_mean, group = VClass, color = VClass)) +
  geom_line() +
  # scale_color_viridis_d() +
  geom_segment(aes(xend = 2021.2, yend = co2_mean), linetype = 2, colour = 'grey') +
  geom_point(size = 2) +
  geom_text(aes(x = 2021.8, label = VClass, size = 8), hjust = 0) +
  labs(title = "Diferent size vehicles emission level",
       subtitle = "2 year average emisson", 
       y = "CO2 (grams / miles)", x = "", caption = "Source: EPA | Grphic: Yifei Liu") +
  theme_ipsum_rc(grid = "Y") +
  theme(legend.position = "none",
        plot.margin = margin(5.5, 160, 5.5, 5.5),
        plot.title = element_text(hjust = 0.5, size = 40),
        plot.subtitle = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        plot.caption = element_text(size = 20)) 

p + transition_reveal(year) + 
  coord_cartesian(clip = 'off') 

animate(p + transition_reveal(year) + 
  coord_cartesian(clip = 'off') , height = 1000, width =2000)

# anim_save("vclass_emission.gif")


```

gift show how emission cross most weight class vehicles decrease. we can see cross different size of vehicles, co2 emission goes down. 

## Most green car brand

Use brand car co2 emission average data, see which brand cars are most green, 

first only use top 10 brand in terms of number of cars on EPA list. 

```{r}
top_emission <- car_emission %>% 
  select(year, make, co2TailpipeGpm) %>%
  group_by(year, make) %>%
  summarize(co2_mean = round(mean(co2TailpipeGpm, na.rm = T), 1),
            total = sum(co2TailpipeGpm, na.rm = T),
            n = n(),
            clean_level = round(1 / co2_mean * 1000000)) %>%
  group_by(year) %>%
  top_n(10, wt = total) %>%
  arrange(desc(clean_level)) %>%
  mutate(ordering = as.double(rev(seq(10:1))) * 1.0,
         make = str_remove(make, "-.*")) # short car companies names
```


```{r}

car_own1 <- read_table("BMW Group owns BMW, Mini, and Rolls-Royce.
Daimler AG owns Mercedes-Benz and Smart.
Fiat Chrysler Automobiles owns Alfa Romeo, Chrysler, Dodge, Fiat, Jeep, Maserati, and Ram.
Ford Motor Co. owns Ford and Lincoln.
General Motors owns Buick, Cadillac, Chevrolet, and GMC.
Honda Motor Co. owns Acura and Honda.
Hyundai Motor Group owns Genesis, Hyundai, and Kia.
Mazda Motor Corp. owns Mazda.
Renault-Nissan-Mitsubishi Alliance owns Infiniti, Mitsubishi, and Nissan.
Subaru Corp. owns Subaru.
Tata Motors owns Jaguar and Land Rover.
Tesla owns Tesla.
Toyota Motor Corp. owns Lexus and Toyota. And it has a stake in Subaru and Suzuki.
Volkswagen Group owns Audi, Bentley, Bugatti, Lamborghini, Porsche, and Volkswagen.
Zhejiang Geely Holding Group (ZGH) owns Lotus, Polestar, and Volvo.", col_names = F) 

car_own2 <- read_table("Acura: Honda Motor Company
Alfa Romeo: Fiat Chrysler Automobiles
Audi: Volkswagen Group
BMW: BMW Group
Bentley: Volkswagen Group
Buick: General Motors
Cadillac: General Motors
Chevrolet: General Motors
Chrysler: Fiat Chrysler Automobiles
Dodge: Fiat Chrysler Automobiles
Fiat: Fiat Chrysler Automobiles
Ford: Ford Motor Co.
GMC: General Motors
Genesis: Hyundai Motor Group
Honda: Honda Motor Co.
Hyundai: Hyundai Motor Group
Infiniti: Renault-Nissan-Mitsubishi Alliance
Jaguar: Tata Motors
Jeep: Fiat Chrysler Automobiles
Kia: Hyundai Motor Group
Land Rover: Tata Motors
Lexus: Toyota Motor Corp.
Lincoln: Ford Motor Co.
Lotus: Zhejiang Geely Holding Group 
Maserati: Fiat Chrysler Automobiles
Mazda: Mazda Motor Corp.
Mercedes-Benz: Daimler AG
Mercury*: Ford Motor Co.
Mini: BMW Group
Mitsubishi: Renault-Nissan-Mitsubishi Alliance
Nissan: Renault-Nissan-Mitsubishi Alliance
Polestar: Zhejiang Geely Holding Group 
Pontiac*: General Motors
Porsche: Volkswagen Group
Ram: Fiat Chrysler Automobiles
Rolls-Royce: BMW Group
Saab: brand owned by Saab AB; assets owned by National Electric Vehicle Sweden
Saturn*: General Motors
Scion*: Toyota Motor Corp.
Smart: Daimler AG
Subaru: Subaru Corp.
Suzuki*: Suzuki Motor Corp. Owns a small stake in Toyota.
Tesla: Tesla Inc.
Toyota: Toyota Motor Corp.
Volkswagen: Volkswagen Group
Volvo: Zhejiang Geely Holding Group", col_names = F)

top_car_make <- top_emission %>% ungroup() %>% count(make)

make_owner1 <- car_own1 %>%
  as_tibble() %>%
  separate(X1, into = c("Own", "Brand"), sep = " owns ") %>% 
  separate_rows(Brand, sep = ",|and|And") %>%
  filter(str_detect(Brand, paste0(top_car_make$make, collapse = "|"))) %>% 
  mutate(Brand = str_match(Brand, paste0(top_car_make$make, collapse = "|"))) %>%
  filter(.$Brand != "Volvo")

make_owner1 <- data.frame(Own = make_owner1$Own,
               make = make_owner1$Brand[,1]) %>%
  add_row(Own = make_owner1$Own,
         make = make_owner1$Own)

make_owner2 <- car_own2 %>%
  as_tibble() %>%
  separate(X1, into = c("make", "Own"), sep = ": ") %>%
  mutate(Own = case_when(make == "Saab" ~ "Saab AB",
                         make == "Suzuki*" ~ "Suzuki Motor Corp",
                         T ~ Own)) %>%
  select(Own, make)

make_owner <- make_owner1 %>%
  full_join(make_owner2, by = c("make", "Own")) %>%
  mutate(make = str_remove(make, "\\*"))
 

top_emission <- top_emission %>%
  left_join(make_owner, by = "make") %>%
  mutate(Own = case_when(make == "Oldsmobile" ~ "General Motors",
                         make == "Plymouth" ~ "Fiat Chrysler Automobiles",
                         make == "Eagle" ~ "Fiat Chrysler Automobiles",
                         T ~ Own))

```



```{r}
p <- top_emission %>%
  #filter(year %% 2 == 0) %>%
  # filter(year == 2020) %>%
  ggplot(aes(ordering, group = make)) +
  geom_tile(aes(y = clean_level / 2,
                height = clean_level,
                width = 0.9,
                fill = Own), alpha = 0.9) +
  geom_text(aes(y = clean_level * 0.95, label = make), size = 5) +
  geom_text(aes(y = clean_level * 1.05, label = clean_level), size = 5) +
  geom_text(aes(x = 1, y = 2700, label = paste0(year)), size = 15, color = "gray45") +
  coord_cartesian(clip = "off", expand = F) +
  scale_fill_viridis_d() +
  coord_flip() +
  labs(title = "Greenest Car Manufactures Company",
       subtitle = "Top 10 Car manufactures in terms of # of cars EPA had tested | 1 ton of CO2 how many miles can you drive",
       x = "", y = "Clean level (unit: miles / ton (CO2))",
       caption = "Source: EPA | Graphic: Yifei Liu")  +
  theme(legend.position = "none",
        text = element_text(family = 'Lato'),
        rect = element_rect(fill = "white"),
        plot.title = element_text(hjust = 0.5, size = 32),
        plot.subtitle = element_text(size = 24),
        plot.caption = element_text(size = 16),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 16),
        axis.title.x = element_text(size = 24),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color = "grey75"),
        panel.grid.minor.x = element_line(color = "grey75"))

animate(p +
  transition_states(year) +
  ease_aes('cubic-in-out') , height = 1000, width = 2000)

anim_save("greenest_companies.gif")


```




Which brand car most economic efficient. 

```{r}

big_epa_cars %>%
  group_by(year, make, electr)

```



Does each Car maker make oil efficient car or not efficient cars?

```{r}

epa_price <- big_epa_cars %>%
  select(year, make, youSaveSpend) %>%
  left_join(make_owner, by = "make") %>%
  mutate(Own = case_when(make == "Oldsmobile" ~ "General Motors",
                         make == "Plymouth" ~ "Fiat Chrysler Automobiles",
                         make == "Eagle" ~ "Fiat Chrysler Automobiles",
                         make == "Isuzu" ~ "General Motors",
                         make == "MINI" ~ "BMW Group",
                         make == "Geo" ~ "General Motors",
                         make == "Lamborghini" ~ "Volkswagen Group",
                         T ~ Own)) %>%
  filter(!is.na(Own)) %>%
  mutate(Own = fct_lump(Own, 19)) %>%
  group_by(year, Own) %>%
  summarize(save_median = median(youSaveSpend)) 

epa_price %>%
  ggplot(aes(year, Own, fill = save_median)) +
  geom_tile() +
  scale_fill_carto_c(palette = "TealRose", direction = -1) +
  labs(title = "Which car cost less money down the road",
       fill = "Meidan save",
       x = "", y = "") +
  guides(fill = guide_colourbar(barheight = 0.3, barwidth = 20, direction = "horizontal", ticks = F)) +
  theme(axis.text.x = element_text(hjust = 1),
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(face = "italic", size = 8),
        axis.ticks = element_blank(),
        axis.title.y = element_text(hjust = 1, size = 7),
        panel.grid = element_blank(),
        legend.position = "bottom")

ggsave("car_save.png", width = 20, height = 15, units = "cm") 

```


How many car models have been tested by EDA

```{r}

e_car <- big_epa_cars %>%
  mutate(year = floor((year + 2) / 6) * 6,
         year = case_when(year == 2022 ~ 2016,
                          T ~ year),
         year = case_when(year == 1986 ~ "1984-1989",
                          year == 1992 ~ "1990-1995",
                          year == 1998 ~ "1996-2001",
                          year == 2004 ~ "2002-2007",
                          year == 2010 ~ "2008-2013",
                          year == 2016 ~ "2014-2020"),
         uses_electricity = case_when(phevHwy > 0 ~ "Hybrid Vehicles",
                                      T ~ uses_electricity),
         uses_electricity = fct_relevel(uses_electricity, c("Fossil Fuel Vehicles", "Electric Vehicles","Hybrid Vehicles"))) %>%
  select(year, uses_electricity) %>%
  group_by(year, uses_electricity) %>%
  count() %>%
  mutate(n = round(n / 100))
  
  
e_car %>%
  ggplot(aes(fill = uses_electricity, values = n)) +
  geom_waffle(color = "white", size = 0.25, nrow = 10, flip = T) +
  facet_wrap(~ year, nrow = 1, strip.position = "bottom") +
  scale_x_discrete() +
  scale_y_continuous(labels = function(x) x * 1000,
                     expand = c(0, 0)) +
  coord_equal() +
  scale_fill_tableau(name = NULL) +
  labs(title = "Beginning of a new area",
       subtitle = "# of vechicle models tested by EPA",
       caption = "Source: EPA | Graphic: Yifei Liu") +
  theme_econodist() +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_line(),
        plot.caption = element_text(hjust = 0),
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = T)) 

# ggsave("e_car.png", width = 20, height = 8, units = "cm")
  
```













