---
title: "Golden age of Television"
author: "Yifei Liu"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(broom)
library(splines)
detach("package:dplyr", unload=TRUE)
library(dplyr)
theme_set(theme_minimal())


tv_ratings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")

```
```{r}
tv_ratings %>%
  filter(title == "Buffy the Vampire Slayer")
```


## Does most shows get better or worse over time?
```{r}
tv_ratings %>%
  count(seasonNumber) %>%
  ggplot(aes(seasonNumber, n)) +
  geom_line()

tv_ratings %>%
  filter(seasonNumber <= 7) %>%
  group_by(seasonNumber) %>%
  summarize(av_rating = mean(av_rating)) %>%
  ggplot(aes(seasonNumber, av_rating)) +
  geom_line()

tv_ratings %>%
  filter(seasonNumber <= 7) %>%
  group_by(seasonNumber) %>%
  summarize(av_rating = mean(av_rating)) %>%
  ggplot(aes(seasonNumber, av_rating)) +
  geom_line()

tv_ratings_full_show <- tv_ratings %>%
  group_by(title) %>%
  mutate(row_number = row_number()) %>%
  filter(1 %in% seasonNumber,
         all(seasonNumber == row_number)) %>%
  arrange(date)


tv_ratings_full_show %>%
  filter(n() >= 5)%>%
  filter(seasonNumber <= 5) %>%
  group_by(seasonNumber) %>%
  summarize(av_rating = mean(av_rating))

tv_ratings_full_show %>%
  filter(n() >= 5)%>%
  filter(seasonNumber <= 5) %>%
  ggplot(aes(factor(seasonNumber), av_rating)) +
  geom_boxplot()

```

we may see in this chart that the average rating of first season is relative low compare to other season. But this could be a data quality issues. We don't have first several season in some tv showes. 
After we filter out the incomplete show, we can see the average rating of differnt season show have simililary rating. 

what are the highest rated seaosn of tv

```{r}
by_show <- tv_ratings %>%
  group_by(title) %>%
  summarize(avg_rating = mean(av_rating),
            sd = sd(av_rating),
            seasons = n()) %>%
  arrange(desc(avg_rating))

most_variable_shows <- by_show %>%
  filter(seasons > 5) %>%
  top_n(6, sd)


tv_ratings %>%
  inner_join(most_variable_shows, by = "title") %>%
  ggplot(aes(seasonNumber, av_rating, color = title)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = function(lim) print(seq(floor(lim[1]), ceiling(lim[2])))) +
  facet_wrap(~ title) +
  theme(legend.position = "none")

```

### Has TV been gettng better or worse over time. 

```{r}
tv_ratings %>%
  group_by(year = year(date)) %>%
  summarize(mean_ratings = mean(av_rating),
            median_rating = median(av_rating),
            sd = sd(av_rating)) 

by_year <- tv_ratings %>%
  group_by(year = 2* (year(date) %/% 2)) %>%
  summarize(mean_ratings = mean(av_rating),
            median_rating = median(av_rating),
            mean_seaosn_1 = mean(av_rating[seasonNumber == 1]),
            avg_season = mean(seasonNumber),
            sd = sd(av_rating),
            observations = n()) 

by_year %>%
  gather(metric, value, -year, - observations, - sd, -avg_season) %>%
  ggplot(aes(year, value, color = metric)) +
  geom_line() +
  geom_point()


tv_ratings %>%
  group_by(year = 5 * (year(date) %/% 5),
           seasonNumber = ifelse(seasonNumber >= 4, "4+", seasonNumber)) %>%
  summarize(mean_rating = mean(av_rating),
            observations = n()) %>%
  ggplot(aes(year, mean_rating, color = seasonNumber)) +
  geom_line() +
  labs(color = "Season",
       y = "Average rating in each 5 year period")


tv_ratings %>%
  ggplot(aes(date, av_rating)) +
  geom_point() +
  geom_smooth(method = "loess")

tv_ratings %>%
  ggplot(aes(date, av_rating)) +
  geom_point() +
  geom_smooth(method = "loess")

tv_ratings %>%
  filter(seasonNumber == 1) %>%
  group_by(year = 2* (year(date) %/% 2)) %>%
  ggplot(aes(year, av_rating, group = year)) +
  geom_boxplot()


tv_ratings %>%
  filter(str_detect(title, "Friends"))


```


### Show survival

If season 1 is good, what's the probability they get a season 2.

Questions:
Whether a show gets a second season depends on the ratings of the first season

```{r}

first_three_seasons <- tv_ratings %>%
  filter(seasonNumber <= 3) %>%
  group_by(title) %>%
  mutate(date_first_season = min(date)) %>%
  ungroup() %>%
  transmute(titleId,
            title,
            date_first_season,
            seasonNumber = paste0("season", seasonNumber),
            av_rating) %>%
  distinct(title, seasonNumber, .keep_all = T) %>%
  spread(seasonNumber, av_rating) %>%
  filter(!is.na(season1)) %>%
  arrange(desc(date_first_season)) %>%
  filter(date_first_season <= "2017-01-01")


glm(!is.na(season2) ~ season1, data = first_three_seasons) %>%
  summary()

```

```{r}

first_three_seasons %>%
  group_by(season1_bin = cut(season1, c(0, 7, 7.5, 8, 8.5, 10)),
           time_bin = ifelse(date_first_season < "2000-01-01", "Before 2000", "After 2000")) %>%
  summarize(had_second_season = mean(!is.na(season2)),
            observations = n()) %>%
  ggplot(aes(season1_bin, had_second_season, group = time_bin, color = time_bin)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = percent_format())
  


```




```{r}


new_data <- crossing(year = 1990:2018,
                     season1 = seq(6, 9))

# assume linear relatiosnhip in year

mod <- first_three_seasons %>%
  mutate(year = year(date_first_season),
         had_season_season = !is.na(season2)) %>%
  glm(had_season_season ~ season1 * year, data = ., family = "binomial")

summary(mod)

mod %>%
  augment(newdata = new_data, type.predict = "response") %>%
  ggplot(aes(year, .fitted, color = factor(season1))) +
  geom_line() +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Probability of getting season 2 has become mroe depend on ratings of season 1",
       color = "Season1 rating",
       x = "year show premiered", 
       y = "Predicted Probability pf getting season 2")




```








