---
title: "median_franschise_powerhouses"
author: "Yifei Liu"
date: "7/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}
library(dplyr)
library(tidyverse)
library(rvest)
library(janitor)
library(ggthemes)
library(RColorBrewer)
library(glue)
library(ggrepel)
library(skimr)
library(ggalluvial)
library(here)
library(cowplot)
theme_set(theme_minimal())

```

load dataset
```{r}
media_franchises <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-02/media_franchises.csv")

```

Take a look at each franchise, see which one have the highest revenue.

```{r}
media_franchises %>%
  group_by(revenue_category) %>%
  summarize(total = sum(revenue)) %>%
  arrange(desc(total))

media_franchises %>%
  group_by(franchise) %>%
  summarize(total = sum(revenue)) %>%
  arrange(desc(total))

media_franchises %>%
  group_by(owners) %>%
  summarize(total = sum(revenue)) %>%
  arrange(desc(total))


media_franchises %>%
  distinct(franchise, .keep_all = T) %>%
  count(owners, sort = T)

franchises <- media_franchises %>%
  group_by(franchise, original_media, year_created, creators, owners) %>%
  summarize(categories = n(),
            total_revenue = sum(revenue),
            #which.max select msot profitbale revenue category
            most_profitable = revenue_category[which.max(revenue)]) %>%
  ungroup()

```

we can see by categories most profit category is merchandise, Licensing & retial. And most valubate media franchises is Pokemon. Walt Disney company own the most valuable franchise in term of # and total $ terms. 


### EDA

#### Franchise-level 

```{r}
franchises %>%
  arrange(desc(revenue))
  


```




```{r}
top_franchise <- franchises %>%
  mutate(franchise = glue("{ franchise } ({ year_created })")) %>%
  top_n(16, total_revenue)

  
  
media_franchises %>%
  mutate(franchise = glue("{ franchise } ({ year_created })")) %>%
  semi_join(top_franchise, by = "franchise") %>%
  mutate(franchise = fct_reorder(franchise, revenue, sum),
         revenue_category = fct_reorder(revenue_category, revenue, sum)) %>%
  ggplot(aes(franchise, revenue)) +
  geom_col(aes(fill = revenue_category)) +
  coord_flip() +
  geom_text(aes(y = total_revenue, 
                label = paste0(scales::dollar(total_revenue, accuracy = 1))), 
                hjust = -0.2, data = top_franchise) +
  scale_fill_viridis_d(direction = -1) +
  scale_y_continuous(labels = scales::dollar_format(suffix = "B")) +
  labs(x = "", y = "Revenue (Billion)",
       title = "Highest Grossing Media Franchises",
       fill = "",
       caption = "Data source: Wikipedia") +
  expand_limits(y = 100) +
  guides(fill = guide_legend(reverse = T)) +
  theme_fivethirtyeight() +
  theme(legend.direction = "vertical",
        legend.position = c(0.8, 0.25),
        legend.background = element_blank(),
        legend.text = element_text(face = "bold"),
        axis.text.y = element_text(face = "bold"),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5))


```

we can see which media franchise is most profitable, which forms of media most profitable. 


take a look at see which owners have more than one franchise. 



```{r}
franchises %>%
  ggplot(aes(year_created, total_revenue)) +
  geom_point(aes(color = original_media, size = total_revenue)) +
  geom_text(aes(label = franchise), check_overlap = T, vjust = 1, hjust = 1) +
  labs(title = "when were the most profitable franchise been created") +
  expand_limits(x = 1910) +
  guides(size = FALSE)

```

which media format make the most amount of money (origin/revenue_categories)

```{r}
media_franchises %>%
  group_by(original_media) %>%
  summarize(revenue = sum(revenue)) %>%
  arrange(desc(revenue))


media_franchises %>%
  group_by(revenue_category) %>%
  summarize(revenue = sum(revenue)) %>%
  arrange(desc(revenue))

# use heatmap




```


```{r}

original_media_revenue_categories <- media_franchises %>%
   mutate(revenue_category = case_when(revenue_category == "Book sales" ~ "Books",
                                      revenue_category == "Box Office" ~ "Movies",
                                      revenue_category == "Comic or Managa" ~ "Manga",
                                      revenue_category == "Home Video/Entertainment" ~ "Home video",
                                      revenue_category == "Merchandise, Licensing & Retail" ~ "Merchandise",
                                      revenue_category == "Video Games/Games" ~ "Games",
                                      T ~ revenue_category)) %>%
  group_by(original_media) %>%
  filter(sum(revenue) >= 45) %>%
  group_by(original_media, revenue_category) %>%
  summarize(revenue = sum(revenue)) %>%
  ungroup() %>%
  mutate(revenue_category = fct_reorder(revenue_category, revenue, sum),
         original_media = fct_reorder(original_media, -revenue, sum))




```



```{r}

original_media_revenue_categories %>% 
  mutate(original_media = fct_rev(original_media),
         revenue_category = fct_rev(revenue_category)) %>%
  ggplot(aes(original_media, revenue_category)) +
  geom_tile(aes(fill = revenue)) +
  coord_flip() +
  labs(x = "",
       y = "",
       fill = "Revenue") +
  scale_fill_viridis_c(direction = -1, labels = scales::dollar) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.key.width = unit(2, "cm"))

```


## Excellent examples from other tidytuesday participants


```{r}

media_clean <- media_franchises %>%
  mutate(original_media = case_when(original_media == "Comic strip"  ~ "Cartoon",
                                    original_media == "Animated film" ~ "Cartoon",
                                    original_media == "Animated series" ~ "Cartoon",
                                    original_media == "Animated cartoon" ~ "Cartoon",
                                    original_media == "Visual novel" ~ "Novel",
                                    original_media == "Anime" ~ "Manga",
                                    T ~ original_media),
         original_media = fct_lump(original_media, 8),
         revenue_category = case_when(revenue_category == "Book sales" ~ "Books",
                                      revenue_category == "Box Office" ~ "Movies",
                                      revenue_category == "Comic or Managa" ~ "Manga",
                                      revenue_category == "Home Video/Entertainment" ~ "Home video",
                                      revenue_category == "Merchandise, Licensing & Retail" ~ "Merchandise",
                                      revenue_category == "Video Games/Games" ~ "Games",
                                      T ~ revenue_category))

media_group <- media_clean %>%
  group_by(original_media, revenue_category, ) %>%
  summarize(total = n(),
            total_revnue = sum(revenue))

meida_lodes <- to_lodes_form(media_group, axes = 1:2)

my_colors <- colorRampPalette(brewer.pal(8, "Dark2"))(17)

meida_lodes %>%
  ggplot(aes(x = x, y = total_revnue, stratum = stratum, alluvium = alluvium,
             fill = stratum, label = stratum)) +
  scale_x_discrete(expand = c(.05, .05), labels = c("Original Format", "Revenue Streams"),
                   position = "top") +
  scale_fill_manual(values = my_colors) +
  geom_flow(width = 1/6) +
  geom_stratum(alpha = .5, width = 2/12, size =  1, fill = NA, color = "grey") +
  geom_text(stat = "stratum", size = 4, family = "Rockwell") +
  labs(x = "", y = "",
       title = "How to make money from different media format") +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text.x = element_text(face = "bold"))
  

media_franchises %>%
  filter(original_media == "Anime")

media_franchises %>%
  mutate(original_media = case_when(original_media == "Comic strip"  ~ "Cartoon",
                                    original_media == "Animated film" ~ "Cartoon",
                                    original_media == "Animated series" ~ "Cartoon",
                                    original_media == "Animated cartoon" ~ "Cartoon",
                                    original_media == "Visual novel" ~ "Novel",
                                    T ~ original_media)) %>%
  count(original_media, sort = T) 

```



```{r}
sorted_rev <- media_franchises %>%
  group_by(franchise) %>%
  mutate(revenue_perc = round(revenue/sum(revenue)*100, 1)) %>% 
  select(franchise, revenue_category, revenue, revenue_perc) %>% 
  arrange(franchise, -revenue_perc) %>%
  mutate(order = row_number())


p <- sorted_rev %>%
  ungroup() %>% 
  mutate(franchise = str_replace(franchise, " / ", "\n")) %>% 
ggplot() +
  geom_col(aes(x = order,
               y = revenue_perc,
               fill = revenue_category),
           # width affects only the biggest revenue, gives warnings
           width = 0.9
           ) +
  facet_wrap(~franchise, ncol = 10) +
  scale_x_reverse() +
  scale_fill_brewer(palette = "Dark2") +
  coord_polar(theta = "y") +
  labs(
    title = "Media Franchise Powerhouses",
    subtitle = "Different revenue streams as percentage of the total revenue.\nThe outer ring shows the largest revenue stream (full circle is 100%)",
    caption = "Source: Wikipedia | Graphic: Yifei Liu"
  ) +
  theme_void() +
  theme(text = element_text(family = "Space Mono"),
    legend.position = "none",
    plot.background = element_rect(fill = "#fff5ba", color = "#fff5ba"),
    plot.margin = margin(0, 50, 0, 50),
    strip.text.x = element_text(size = 1),
    plot.title = element_text(size = 11, hjust = 0.5,
                              margin = margin(30, 0, 5, 0)),
    plot.subtitle = element_text(size = 7, hjust = 0.5,
                              margin = margin(0, 0, 30, 0)),
    plot.caption = element_text(size = 6, hjust = 0.5,
                                margin = margin(30, 0, 30, 0)))

# custom "legend" plot
l <- tribble(
  ~category, ~x, ~y,
  "Video Games/Games", 1.5,  2,
  "Box Office", 2.15,  2,
  "Home Video/Entertainment", 3.65, 2,
  "Music", 4, 2,
  "TV", 0.5, 1,
  "Book Sales", 1.15, 1,              
  "Merchandise, Licensing & Retail", 3.1, 1,
  "Comic or Manga", 4, 1
  ) %>% 
  ggplot(aes(label = category, x = x, y = y,
             color = category)) +
  # geom_text(family = "Space Mono Bold",
  #           hjust = 1, size = 1.5) +
  geom_label(aes(fill = category),
             label.r = unit(0, "lines"),
             label.padding = unit(0.05, "lines"),
             color = "#fff5ba",
             hjust = 1, size = 1.5) +
  scale_fill_brewer(palette = "Dark2") +
  coord_fixed(ratio = 0.2, xlim = c(0, 4), ylim = c(0, 2.5)) +
  theme_void() +
  theme(text = element_text(family = "Space Mono"),
    legend.position = "none",
    plot.background = element_rect(fill = "#fff5ba", color = "#fff5ba")
  )

ggdraw() +
  draw_plot(p) +
  draw_plot(l, x = 0.35, y = 0.11, width = 0.52, height = 0.1) 




```



```{r}
media_franchises %>%
  group_by(franchise) %>%
  mutate(revenue_pct = round(revenue / sum(revenue) * 100, 1)) %>%
  select(franchise, revenue_category, revenue, revenue_pct) %>%
  arrange(franchise, - revenue_pct) %>%
  mutate(order = row_number())
  
sorted_rev %>%
  ungroup() %>%
  mutate(franchises = str_replace(franchise, " / ", "\n")) %>%
  ggplot() +
  geom_col(aes(x = order, y = revenue_perc, fill = revenue_category), width = 1) +
  facet_wrap(~ franchises, ncol = 10) +
  scale_x_reverse() +
  scale_fill_brewer(palette = "Dark2") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Media Franchises Powerhouse",
       subtitle = "Differet revneue streams as percentage of the total revneu. \nThe outer ring shows the largest revneue stream(full circle is 100%)",
       caption = "Data source: Wikipedia | Graphic: Yifei Liu") +
  theme(legend.position = "none",
        text = element_text(family = "Space Mono"),
        plot.background = element_rect(fill = "#fff5ba", color = "#fff5ba"),
    plot.margin = margin(0, 50, 0, 50),
    strip.text.x = element_text(size = 1),
    plot.title = element_text(size = 11, hjust = 0.5,
                              margin = margin(30, 0, 5, 0)),
    plot.subtitle = element_text(size = 7, hjust = 0.5,
                              margin = margin(0, 0, 30, 0)),
    plot.caption = element_text(size = 6, hjust = 0.5,
                                margin = margin(30, 0, 30, 0)))

```




















