---
title: "R4DsMembership"
author: "Yifei Liu"
date: "8/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and pacakges

```{r}
library(tidyverse)
library(skimr)
library(lubridate)
library(ggthemes)
library(ggTimeSeries)
library(viridis)
theme_set(theme_minimal())

r4ds_members <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-16/r4ds_members.csv") %>%
  mutate(date = ymd(date),
         year = factor(year(date)),
         month = factor(month(date, label = T), levels = month.abb),
         weekday = factor(wday(date, label = T)))


```


### Inspect data

```{r}
skim(r4ds_members)

```

### EDA

let's take a look see how many people join R4DS every day

```{r}
r4ds_members %>%
  mutate(daily_in = total_membership - lag(total_membership, default = NA)) %>%
  ggplot(aes(daily_in)) +
  geom_histogram(binwidth = 10)

r4ds_members %>%
  mutate(daily_in = total_membership - lag(total_membership, default = NA)) %>%
  ggplot(aes(date, daily_in)) +
  geom_line()

r4ds_members %>%
  gather(key = "key", value = "value", -date) %>%
  ggplot(aes(date, value)) +
  geom_line() +
  facet_wrap(~ key, scales = "free_y")


```

use lubridate pacakge to analyze this time series dataset.


```{r}
r4ds_members %>%
  mutate(wday = wday(date, label = T)) %>%
  select(wday, daily_active_members, daily_members_posting_messages) %>%
  gather(key = "key", value = "value", -wday) %>%
  ggplot(aes(wday)) +
  geom_bar() +
  facet_wrap(~ key, scales = "free_y")

ggplot(r4ds_members, aes(date, daily_active_members))+
  geom_line(lwd = 1)+
  facet_wrap(~year, ncol = 1, scales = "free")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  theme_minimal()


r4ds_members %>%
  select(daily_active_members, daily_members_posting_messages, date, year) %>%
  gather(key = "key", value = "value", daily_active_members, daily_members_posting_messages) %>%
  mutate(key = str_to_title(str_replace_all(key, "_", " "))) %>%
  ggplot(aes(date, value)) +
  geom_area(aes(fill = key), alpha = 0.9) +
  facet_wrap(~year, ncol = 1, scales = "free")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  scale_fill_brewer(type = "Seq", palette = "Blues") +
  labs(x = "", y = "Number of Memebers", fill = NULL) +
  theme(legend.position = "top",
        text = element_text(family = "Space Mono"))

r4ds_members %>%
  select(weekly_active_members, weekly_members_posting_messages, date, year) %>%
  gather(key = "key", value = "value", weekly_active_members, weekly_members_posting_messages) %>%
  mutate(key = str_to_title(str_replace_all(key, "_", " "))) %>%
  ggplot(aes(date, value)) +
  geom_area(aes(fill = key), alpha = 0.9) +
  facet_wrap(~year, ncol = 1, scales = "free")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  scale_fill_brewer(type = "Seq", palette = "Reds") +
  labs(x = "", y = "Number of Memebers", fill = NULL) +
  theme(legend.position = "top",
        text = element_text(family = "Space Mono"))



```

### Excellient work from other #tidytuesday participant 


#### [Calender heatmap](https://github.com/NewMirai/Tidytuesday/blob/master/tidytuesdayR4DS/plot.r)



```{r}
r4ds_members %>%
  ggplot_calendar_heatmap('date', 'daily_active_members', monthBorderSize = 1.5, monthBorderColour = "black") +
  scale_fill_viridis(option = "D") +
  theme_minimal() +
  facet_wrap(~year, ncol = 1, strip.position = "right") +
  theme(text = element_text(family = "Space Mono"),
        panel.spacing = unit(5, "lines"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = NA, fill="black"),
        plot.background = element_rect(fill="black"),
        legend.position = "bottom",
        axis.text.x = element_text(size = 12,colour ="#bfbfbf",vjust = -2),
        axis.text.y = element_text(size=12,colour = "#bfbfbf"),
        strip.text = element_text(size=14,colour="#bfbfbf"),
        legend.text = element_text(colour = "#bfbfbf",size=8),
        legend.key.width = unit(1.2, "cm"),
        legend.title = element_text(colour = "#bfbfbf",vjust = .9,size=14),
        plot.title = element_text(size=20,colour="#bfbfbf",hjust = .5),
        plot.subtitle = element_text(size=16,colour = "#bfbfbf",hjust = .5),
        plot.caption = element_text(colour = "#bfbfbf"))+
  labs(y='',
       fill="Daily active members",
       title = "Calendar heatmap of the active members",
       subtitle = "Made with the ggTimeSeries package for super easy calendar heatmap\n",
       caption = "#Tidy tuesday | Source: R4DS Slack | @alangel12407606")


```

my replicated without look at the code.

```{r}
r4ds_members %>%
  ggplot_calendar_heatmap('date', 'daily_members_posting_messages', monthBorderSize = 1.5, monthBorderColour = "black") +
  scale_fill_viridis() +
  facet_wrap(~year, strip.position = "right", ncol = 1) +
  labs(title = "Calender heatmap of most active post",
       subtitle = "with ggTimeSeries package",
       x = "", y = "", fill = "Daily Members Posting Messages", 
       caption = "#Tidytuesday | DataSource: R4DS Slack | @yifei liu") +
  theme(text = element_text(family = "Space Mono", color = "#bfbfbf"),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = NA, fill="black"),
        plot.background = element_rect(fill="black"),
        legend.key.width = unit(1.4, "cm"))



```

I think this is great way to visualize these type of calender activities, scuh as flight data, electricity usage data etc. I will definitely use these more in the future. 
















