---
title: "school diversity"
author: "Yifei Liu"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(tigris)
detach("package:dplyr", unload = TRUE)
library(dplyr)
library(sf)
library(rcartocolor)
library(tidytext)
options(tigris_class = "sf")
library(showtext)
font_add_google("Lato", regular.wt = 300, bold.wt = 700)

# custome ggplot theme
custom_theme <- theme_minimal() +
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 28),
        plot.subtitle = element_text(hjust = 0.5, size = 16))
theme_set(custom_theme)

# different datasets

```


```{r}
school_diversity <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-24/school_diversity.csv") # schoold diverse


income_variables <- paste0("B19013", c("A", "B", "C", "D", "G","I"), "_001")

income_table <- tibble(income_variables,
                       race =c("White", "Black", "AIAN", "Asian", "Multi", "Hispanic"))

states_income <- get_acs(geography = "state",
                      variables = income_variables,
                      geometry = T,
                      shift_geo = T)


states_income <- states_income %>%
  left_join(income_table, by = c("variable" = "income_variables")) %>%
  select(-variable) %>%
  rename(median_income = estimate) %>%
  mutate(state = state.abb[match(NAME, state.name)]) %>%
  select(state, median_income, moe, race)

```

This dataset only contain elementary shcools. 

# EDA

```{r}
skimr::skim(school_diversity)

```


By reading the dataset description on github and skim the dataset. We know the dataset contain info about *Elementary School District diversity* from two time period, 94/95 and 16/17. Most variables are complete. Varibale such as _int_group_ and _variance_, which measure level of integration and variance ratio are both missing for *90%* of the dataset. Miss same values in _d_locale_txt_. 

Now we can try to answer some basic questions such as what's the average student population size in different state, does schoold diversity change during last 20 years. etc.

```{r}

# Type of School district
school_diversity %>%
  group_by(SCHOOL_YEAR) %>%
  filter(!is.na(d_Locale_Txt)) %>%
  count(d_Locale_Txt, sort = T)

# How many students live in those type of city
school_diversity %>%
  filter(!is.na(d_Locale_Txt)) %>%
  group_by(SCHOOL_YEAR, d_Locale_Txt) %>%
  summarize(pop = sum(Total))

# How many school districts are diverse 

school_diversity %>%
  group_by(SCHOOL_YEAR) %>%
  count(diverse, sort = T)

# How many race people live in different state

school_diversity %>%
  gather(AIAN:Multi, key = "race", value = "pct") %>%
  group_by(SCHOOL_YEAR, ST, race) %>%
  summarize(race_median = median(pct, na.rm = T))

# which state have most school districts
school_diversity %>%
  mutate(ST = state.name[match(ST, state.abb)]) %>%
  group_by(SCHOOL_YEAR, ST) %>%
  count() %>%
  ungroup() %>%
  mutate(ST = reorder_within(ST, n, SCHOOL_YEAR)) %>%
  group_by(SCHOOL_YEAR) %>%
  top_n(n = 12, wt = n) %>%
  ggplot(aes(ST, n, fill = ST)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_wrap(~ SCHOOL_YEAR, scales = "free_y") +
  scale_fill_tableau()
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(x = "", y = "", title = "Which State have the most school district")

# which state have most students

school_diversity %>%
  group_by(SCHOOL_YEAR, ST) %>%
  count() %>%
  ungroup() %>%
  mutate(ST = reorder_within(ST, n, SCHOOL_YEAR)) %>%
  group_by(SCHOOL_YEAR) %>%
  top_n(n = 12, wt = n) %>%
  ggplot(aes(ST, n, fill = ST)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_wrap(~ SCHOOL_YEAR, scales = "free_y") +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(x = "", y = "", title = "Which State have the most school district")

# average school size in different state

school_diversity %>%
  group_by(ST, SCHOOL_YEAR) %>%
  summarize(pop_size = mean(Total, na.rm = T))

# total state student population

school_diversity %>%
  group_by(ST, SCHOOL_YEAR) %>%
  summarize(pop = sum(Total, na.rm = T))


```

```{r}

school_diversity %>%
  group_by(SCHOOL_YEAR, ST) %>%
  summarize(pop = sum(Total)/ 1000000) %>%
  ungroup() %>%
  mutate(ST = reorder_within(ST, pop, SCHOOL_YEAR)) %>%
  group_by(SCHOOL_YEAR) %>%
  top_n(n = 12, wt = pop) %>%
  ggplot(aes(ST, pop, fill = ST)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_wrap(~ SCHOOL_YEAR, scales = "free_y") +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(x = "", y = "", title = "Which state have most students",
       subtitle = "Elementary School population, population unit: million")



```




```{r}
school_diversity %>%
  group_by(SCHOOL_YEAR) %>%
  count(diverse) %>%
  ggplot(aes(diverse, n)) +
  geom_col(aes(fill = diverse)) +
  coord_flip() +
  scale_fill_carto_d() +
  facet_wrap(~ SCHOOL_YEAR) +
  labs(x = "", y = "",
       title = "School become more diverse",
       subtitle = "Elementary Shcool district year: 94/95 and 16/17") +
  theme(legend.position = "bottom")

```

in this chart we can see school become more diverse during the last 20 years, as expected. 



Next we can take a look at how many kids inside U.S. school district live in a diverse, undiverse and extremely undiverse school.


```{r}

school_diversity %>%
  group_by(SCHOOL_YEAR, diverse) %>%
  summarize(pop = sum(Total) / 1000000) %>%
  ggplot(aes(diverse, pop)) +
  geom_col(aes(fill = diverse)) +
  coord_flip() +
  scale_fill_carto_d() +
  facet_wrap(~ SCHOOL_YEAR) +
  labs(x = "", y = "",
       title = "How many students live in a diverse enviornment",
       subtitle = "Elementary Shcool population year: 94/95 and 16/17, in millions") +
  theme(legend.position = "bottom")

```



now put ever student # on map see, which state 

```{r}

race_pop <- school_diversity %>%
  gather(AIAN:Multi, key = "race", value = "pct") %>%
  fill(pct) %>% 
  mutate(pop = pct / 100 * Total / 1000000) %>%
  group_by(ST, SCHOOL_YEAR, race) %>%
  summarize(pop = sum(pop)) 


race_map <- left_join(states_income, race_pop, by = c("state" = "ST", "race"))

race_map %>%
  filter(SCHOOL_YEAR == "2016-2017") %>%
  ggplot() +
  geom_sf(aes(fill = pop), color = NA) +
  scale_fill_viridis_c(direction = -1) +
  facet_wrap(~ race) +
  guides(fill = guide_colourbar(barheight = 0.3, barwidth = 20, direction = "horizontal", ticks = F)) +
  labs(title = "minority student concentrate in Costal cities",
       subtitle = "Year: 16/17, population unit: million") +
  theme(legend.position = "bottom")
  

```



```{r}





```












