---
title: 'TidyTuesday: Simpsons Guest Stars'
author: "Yifei Liu"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(tidytext)
library(ggtext)
library(hrbrthemes)


custome_theme <- function (econ_text_col = "#3b454a", econ_plot_bg_col = "#d7e6ee", 
  econ_grid_col = "#bbcad2", econ_font = "Roboto-Regular", 
  light_font = "Roboto-Light", bold_font = "Roboto-Bold") 
{
  theme_minimal(base_family = econ_font) + 
    theme(plot.title = element_text(family = bold_font), 
    plot.subtitle = element_text(family = light_font, size = 12, 
      margin = margin(t = -5)), plot.caption = element_text(family = light_font, 
      size = 10, colour = econ_text_col, lineheight = 1.1))
}

theme_set(custome_theme())


simpsons <- readr::read_delim("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-27/simpsons-guests.csv", delim = "|", quote = "") %>%
  mutate(self = str_detect(role, "self|selves"),
         season = parse_number(season))

dialogue <- read_csv("/Users/yifeiliu/Documents/R/data/EDA/tidytuesday/simpsons_dataset.csv") %>%
  select(role = raw_character_text,
            lines = spoken_words) # ref 1. first 27 seasons of simpsons 



```


Basic EDA


```{r}
simpsons %>%
  filter(self) %>%
  count(guest_star, sort = T) %>%
  filter(n > 1) %>%
  mutate(guest_star = fct_reorder(guest_star, n)) %>%
  ggplot(aes(guest_star, n)) +
  geom_col() +
  coord_flip() +
  labs(x = "", y = "",
       title = "Who has played themeselves in multiple Simpsons episodes")
  

```


```{r}
simpsons %>%
  separate_rows(role, sep = ";\\s+") %>%
  add_count(role) %>%
  filter(n >= 10) %>%
  count(season, role) %>%
  mutate(role = fct_reorder(role, -n, sum)) %>%
  ggplot(aes(season, n)) +
  geom_col() +
  facet_wrap(~ role, ncol = 1)

```



### Bringing in Simpsons dialogue

```{r}
dialogue %>%
  count(role, sort = TRUE) %>%
  semi_join(simpsons, by = "role")

guests_processed <- simpsons %>%
  separate_rows(role, sep = ";\\s+") %>%
  mutate(role = ifelse(self, guest_star, role),
         role = case_when(role == "Edna Krabappel" ~ "Edna Krabappel-Flanders",
                          role == "Stephen Hawking" ~ "Dr. Stephen Hawking",
                          T ~ role))

guest_summarized <- guests_processed %>%
  filter(season <= 27) %>%
  group_by(guest_star, role, self) %>%
  summarize(nb_spisodes = n(),
            first_season = min(season),
            last_season = max(season)) %>%
  arrange(desc(nb_spisodes)) %>%
  group_by(role) %>%
  filter(n() == 1) %>%
  ungroup() %>%
  filter(!is.na(role))
 
dialogue_summarized <- dialogue %>%
  group_by(role) %>%
  summarize(nb_lines = n(),
            random_line = sample(lines, 1)) %>%
  arrange(desc(nb_lines))


guest_roles <- guest_summarized %>%
  inner_join(dialogue_summarized, by = "role") %>%
  mutate(lines_per_episode = round(nb_lines / nb_spisodes, digits = 3))

guest_roles %>%
  mutate(self = ifelse(self, "Playing themselves", "Playing a Character")) %>%
  ggplot(aes(lines_per_episode)) +
  geom_histogram(binwidth = 2, center = 1) +
  facet_wrap(~ self, ncol = 1) +
  labs(x = "Average # of line per episode",
       title = "Most guest star, espeically those playing themselves, have relatively few lines per episode")



```

### dialogue text analysis

```{r}
role_words <- dialogue %>%
  filter(!is.na(lines),
         !is.na(role)) %>%
  mutate(line_number = row_number()) %>%
  unnest_tokens(word, lines) %>%
  anti_join(stop_words, by = "word") %>%
  distinct(role, line_number, word) %>%
  count(role, word, sort = T)

role_word_tf_idf <- role_words %>%
  group_by(role) %>%
  mutate(total_words = sum(n)) %>%
  ungroup() %>%
  bind_tf_idf(word, role, n) %>%
  arrange(desc(tf_idf))
  
role_word_tf_idf %>%
  filter(total_words >= 500)  %>%
  distinct(role, .keep_all = TRUE) %>%
  mutate(role_word = paste(role, ": ", word)) %>%
  head(20) %>%
  mutate(role_word = fct_reorder(role_word, tf_idf)) %>%
  ggplot(aes(role_word, tf_idf)) +
  geom_col() +
  coord_flip() +
  labs(title = "Using TF-IDF as a Simpsons catchphrase detector",
             subtitle = "Only the 53 cahracters that speak at least 500 words",
             x = "", y = "TF-IDF")
  

```

```{r}
guest_summarized %>%
  filter(nb_spisodes > 1) %>%
  inner_join(role_word_tf_idf, by = "role") %>%
  distinct(role, .keep_all = T)


```

### Other people analysis

1. [gkaramanis](http://bit.ly/tidytue35)

```{r}
sim <- simpsons %>%
  # style for highlighted guest stars
  mutate(
    guest_star = case_when(
      str_detect(tolower(role), "himself|herself|themselves") ~ paste0("<span style='color:black'>", guest_star, "</span>"),
      # style for others
      TRUE ~ paste0("<span style='color:#D1B271'>", guest_star, "</span>")
    )
  ) %>%
  # keep unique combinations
  distinct(guest_star, role)

all_guests <- paste(sim$guest_star, collapse = " • ")

ggplot() +
  # set color to NA to hide the box (it hides unstyled text, too!)
  geom_textbox(aes(label = all_guests, 1, 2), 
               color = NA, fill = NA,
               height = 0.9, width = 0.845,
               box.padding = unit(c(0, 0, 0, 0), "pt"),
               hjust = 0.5,
               family = "Akura Popo") +
    labs(
      title = "The Simpsons Guest Stars",
      subtitle = "Unique guest-role combinations, in order of first appearance. Highlighted are the guest voices that played themselves.",
      caption = "Source: Wikipedia via Andrew Collier | Graphic: Georgios Karamanis"
    ) +
    lims(x = c(0, 2), y = c(0, 2)) +
    theme_void(base_family = "Akura Popo") +
  theme(
      legend.position = "none",
      plot.background = element_rect(fill = "#FED90F", color = "#FED90F"),
      plot.title = element_text(size = 32, hjust = 0.5, margin = margin(40, 0, 0, 0)),
      plot.subtitle = element_text(size = 24, hjust = 0.5),
      plot.caption = element_text(size = 12, hjust = 0.9,
                                  margin = margin(30, 30, 40, 0))
    )



```

I try to replicate his code

```{r}

sim <- simpsons %>%
  mutate(guest_star = case_when(
    str_detect(tolower(role), "self|selves") ~ paste0("<span style='color:black'>", guest_star, "</span>"),
    T ~ paste0("<span style='color:#D1B271'>", guest_star, "</span>")
  )) %>%
  distinct(role, guest_star)

all_guests <- paste(sim$guest_star, collapse = " • ")

ggplot() +
  # set color to NA to hide the box (it hides unstyled text, too!)
  geom_textbox(aes(label = all_guests, 1, 2), 
               color = NA, fill = NA,
               height = 0.9, width = 0.845,
               box.padding = unit(c(0, 0, 0, 0), "pt"),
               hjust = 0.5,
               family = "Akura Popo") +
    labs(
      title = "Simpsons guest star",
       subtitle = "30 Seasons, Guest starts appear as themselves mark as black",
       caption = "Data Source: WikiPedia | Graphic: Yifei Liu"
    ) +
    lims(x = c(0, 2), y = c(0, 2)) +
    theme_void(base_family = "Akura Popo") +
  theme(
      legend.position = "none",
      plot.background = element_rect(fill = "#FED90F", color = "#FED90F"),
      plot.title = element_text(size = 32, hjust = 0.5, margin = margin(40, 0, 0, 0)),
      plot.subtitle = element_text(size = 24, hjust = 0.5),
      plot.caption = element_text(size = 12, hjust = 0.9,
                                  margin = margin(30, 30, 40, 0)))

  
```

What kind of graphic is this: communicate to customer. 


2. [Simpsons' Guest Appearances are growing](https://evamaerey.github.io/tidytuesday_walk_through/simpsons.html#1)

Try to "hand drawing this plot" means wan the axes and the lines connecting the points to look hand drawn, So geom_line + jitter

```{r}
clean_data <- simpsons %>%
  filter(!is.na(season)) %>%
  group_by(season) %>%
  summarize(guest_per_episode = n_distinct(guest_star) / n_distinct(number))


```

_spline()_ to connect points

He uses spline function, which makes a smooth connection between points, rather than sharp, as a human might do in a rush. 

```{r}
spline_int <- as.data.frame(spline(x = clean_data$season,
                     y = clean_data$guest_per_episode,
                     method = "natural")) 


```

Hand draw axes with position_jitter()

```{r}
drawn_x_axis <- data.frame(x = seq(from = 0, to = 30, by = 1),
                           y = rep(0, 31))

drawn_y_axis <- data.frame(x = rep(0, 21),
                           y = seq(from = 0, to = 4, by = 0.2))

```

now plot

```{r}
sims_plot <- clean_data %>% 
  ggplot(aes(x = season, y = guest_per_episode)) +
  geom_point(size = 4, color = "snow3") +
  geom_line(data = spline_int, 
            mapping = aes(x= x, y = y),
            color = "snow3", size = 0.8,
            position = position_jitter(w = 0.05, h = 0.05)) +
  geom_line(data = drawn_x_axis,
            mapping = aes(x = x, y = y),
            color = "snow3", size = 0.8,
            position = position_jitter(w = 0.0, h = 0.05)) +
  geom_line(data = drawn_y_axis,
            mapping = aes(x = x, y = y),
            color = "snow3", size = 0.8,
            position = position_jitter(w = 0.05, h = 0.0)) +
  labs(y = "Guest Per Season", x = "season",
       title = "The number of guest stars in \nthe Simpsons is increasing",
       caption = "Data Source: Wikipedia | Graphic: Yifei Liu") +
  lims(x = c(-0.1, 30.5), y = c(-0.1, 4.3)) +
  theme_void(base_family = "Akura Popo") +
  theme(
    axis.text = element_text(size = 15, color = "snow3"),
    axis.title = element_text(size = 20, color = "snow3"),
    plot.title = element_text(size = 25, face = "bold", color = "snow3", hjust = .5),
    plot.caption = element_text(color = "snow3"),
    plot.margin = margin(1, 3, 1, 1, "cm"),
    axis.title.y = element_text(angle = 90)
  )


```

Now add background

```{r}

simpsons_gif_url <- "https://raw.githubusercontent.com/GilHenriques/TidyTuesdays/master/2019-08-27%20Simpsons%20guest%20stars/chalkboard_simpsons.gif"

sims_plot %>%
  ggimage::ggbackground(
    background = simpsons_gif_url,
    by = "height"
  )

```



Reference: 
1. [Dialogue Lines of The Simpsons](https://www.kaggle.com/pierremegret/dialogue-lines-of-the-simpsons/downloads/dialogue-lines-of-the-simpsons.zip/1)






