---
title: "WIne Rating"
author: "Yifei Liu"
date: "6/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(Matrix)
library(lubridate)
library(tigris)
library(doMC)
registerDoMC(cores = 4)
library(widyr)
library(glmnet)
library(scales)
library(broom)
library(tidytext)
library(maps)
theme_set(theme_minimal())

wine_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-28/winemag-data-130k-v2.csv")  %>%
  mutate(date = str_extract_all(title, "[1-2][9-9][0-9][0-9]")) %>%
  replace_na(list(date = NA)) %>%
  unnest() %>%
  group_by(X1) %>%
  distinct() %>%
  ungroup() %>%
  # filter out item with no price, because price is a powerful predictor for wine points. I want to predict wine point, so just filter no price out. 
  filter(!is.na(price)) %>%
  select(-X1) %>%
  mutate(wine_id = row_number())

```
Not all title have date info, and most wine produce in 20 and 21 century. 

```{r}
wine_ratings
skim(wine_ratings)

wine_ratings %>%
  count(date)

wine_ratings %>%
  count(X1, sort = T)


```

look at all variables, we can say we have relative complete varibale in country, descroption, designation, province, region 1, taster_name, twitter_handle, title, variety, winery, point, price info

We can ask serveral questions base on this dataset such as

1. which year we have the highest point wine. (ggplot + geom_line)
2. which country produce highest quality wine
3. which word in description provided by taster associated with high points / prices
4. which country / provience produce most wine. 

### EDA

see which country, provience, region_1 produce most wine.

```{r}
wine_ratings %>%
  mutate(country = fct_lump(country, n = 10)) %>%
  count(country, sort = T) %>%
  top_n(11) %>%
  mutate(country = fct_reorder(country, n)) %>%
  ggplot(aes(country, n, fill = country)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "",
       y = "",
       title = "Which country produce most wine",
       subtitle = "Points above 80, top 10 grape type count for more than 50% wines")
  
wine_ratings %>%
  count(variety, sort = T) %>%
  top_n(11) %>%
  mutate(variety = fct_reorder(variety, n)) %>%
  ggplot(aes(variety, n, fill = variety)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "",
       y = "",
       title = "Which grape type is most common",
       subtitle = "Points above 80, top 10 countries count for more than 95% wines")
  


```

Which year wine we have in the dataset. 

```{r}
wine_ratings %>%
  count(date) %>%
  ggplot(aes(date, n)) +
  geom_area(fill = "red", alpha = 0.5) +
  labs(x = "",
       y = "# of wines",
       title = "Most of wines come after 2000")

```



Top 10 country, US France, Itlay.. produce more than 95% of wine. And top 0 provience count for more than 8% of wine production. 



let's plot relationship between wine price and point, (be careful since this dataset only include wine which points above 80)


```{r}
wine_ratings %>%
  ggplot(aes(points, price, group = as.factor(points))) +
  geom_boxplot() +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) 


```

we can see a general upward trend in this dataset. as rating of wine incrase, wine price also increase. 

Which country's wine have the highest rating. (medium)

```{r}
wine_ratings %>%
  mutate(country = fct_lump(country, 9),
         country = fct_reorder(country, points)) %>%
  ggplot(aes(country, points)) +
  geom_boxplot(aes(fill = country, color = country), alpha = 0.5, show.legend = F) +
  coord_flip() +
  labs(x = "",
       y = "",
       title = "Which country produce the highest rating wine",
       subtitle = "measurement method use median")

```
Country can also been consider a variable to predict wien quality. 

Now we can build a linear regression model base on taster name, year, country. 

```{r}

model <- wine_ratings %>%
  replace_na(list(taster_name = "Missing", country = "Missing")) %>%
  mutate(country = fct_relevel(fct_lump(country, 7), "US"),
         taster_name = fct_relevel(fct_lump(taster_name, 6), "Missing")) %>%
  lm(points ~ log2(price) + country + date + taster_name, data = .) 

model %>%
  tidy(conf.int = T) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = str_replace(term, "country", "Country: "),
         term = str_replace(term, "taster_name", "Taster: "),
         term = fct_reorder(term, estimate)) %>%
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  labs(x = "",
       y = "")
```

```{r}
model %>%
  augment(date = wine_ratings) %>%
  ggplot(aes(.fitted, points)) +
  geom_point(alpha = 0.1)
```

```{r}
tidy(anova(model)) %>%
  mutate(sumsq / sum(sumsq))
```

base on this anova table, we can say 37& of variance in points can be explain by log2(price), the rest is very minimal, can be neglect. 



let's do some text analysis, which descroption have associated with higher wine rating. 

```{r}
rating_price <- wine_ratings %>%
  select(wine_id, description, points, price)

rate_word <- rating_price %>%
  unnest_tokens(word, description) %>%
  filter(!str_detect(word, "^[0-9]")) %>%
  anti_join(stop_words, by = "word")

rate_word %>%
  count(word, sort = T)
```
After some google artciles, [How to Describe Your Favorite Wine in 6 Words or Less](https://discover.wineaccess.com/describe-favorite-wine-6-words-less/) which describe what word to
describe wine, such as acidity, oak, tanin, fruity, body, e.g. light-bodied, full bodied. 

Now I assume wine produce from one region usually been describe by one of these words, and I can use these describe word to tell where this wine come from. 


I can assume the description word associated with wine quality, and grape type. 


### Lasso regression on words in descroption

```{r}
cus_stop <- stop_words %>%
  filter(word != "not") %>%
  add_row(word = c("wine", "drink"))
  

wine_rating_words <- wine_ratings %>%
  unnest_tokens(word, description) %>%
  filter(str_detect(word, "[a-z]")) %>%
  anti_join(cus_stop, by = "word") %>%
  filter(str_detect(word, "[a-z]"))

# which word most common in descrption
wine_rating_words %>% 
  count(word, sort = T) %>%
  head(20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  ggplot(aes(word, n, fill = word)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "",
       y = "")
  
# counting and correlating pairs of words 

wine_words_filtered <- wine_rating_words %>%
  distinct(wine_id, word) %>%
  add_count(word) %>%
  filter(n > 100)

wine_words_filtered %>%
  pairwise_cor(word, wine_id, sort = T)
  
```


```{r}
wine_word_matrix <- wine_words_filtered %>%
  cast_sparse(wine_id, word)

wine_id <- as.integer(rownames(wine_word_matrix)) 
scores <- wine_ratings$points[wine_id] 

# add prices as a predictors

wine_word_matrix_extra <- cbind(wine_word_matrix, log_price = log2(wine_ratings$price[wine_id]))

cv_glmnet_model <- cv.glmnet(wine_word_matrix_extra, scores, parallel = T)

plot(cv_glmnet_model)

```

use cv.glmnet to test how many variable we start overfitting our model. As we increase of variable, from 300 to 2300 we still not overfitting our model. We still lower return as # increase.  


```{r}
lexicon <- cv_glmnet_model$glmnet.fit %>%
  tidy() %>%
  filter(lambda == cv_glmnet_model$lambda.1se,
         term != "(Intercept)",
         term != "log_price")

lexicon %>%
  arrange(estimate) %>%
  group_by(direction = ifelse(estimate > 0, "Positive", "Negative")) %>%
  top_n(10, abs(estimate)) %>%
  ungroup() %>%
  mutate(term = fct_reorder(term, estimate)) %>%
  ggplot(aes(term, estimate, fill = direction)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "",
       title = "What words are predictive of a wine's score?")


```

```{r}
wine_rating_words %>%
  filter(wine_id %in% sample(unique(wine_id), 6)) %>%
  distinct(word, title, points) %>%
  select(word, title, points) %>%
  mutate(wine = paste0(str_trunc(title, 40), " (", points, ")")) %>%
  inner_join(lexicon, by = c("word" = "term")) %>%
  mutate(word = fct_reorder(word, estimate, mean)) %>%
  ggplot(aes(word, estimate, fill = estimate > 0)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_wrap(~ wine, scales = "free_y") +
  labs(x = "",
       y = "",
       title = "How a lasso regression would predict each wine's score",
       subtitle = "Using a lasso regression with an extra term for price")

```




###  what is glmnet?

```{r}
cv_glmnet_model$glmnet.fit %>%
  tidy() %>%
  filter(term %in% c("rich", "black", "simple", "vineyard", "concentrated")) %>%
  ggplot(aes(lambda, estimate, color = term)) +
  geom_line() +
  scale_x_log10() +
  geom_hline(lty = 2, yintercept = 0)

cv_glmnet_model$glmnet.fit %>%
  tidy() %>% 
  count(lambda) %>%
  ggplot(aes(lambda, n)) +
  geom_line() +
  scale_x_log10()
```





