---
title: "Analyze women's world cup"
author: "Yifei Liu"
date: "7/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(countrycode)
library(lubridate)
library(rvest)
library(StatsBombR) # devtools::install_github("statsbomb/StatsBombR")
library(gganimate)
theme_set(theme_minimal())

wwc_outcomes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv")
squads <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
codes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")


outcomes <- wwc_outcomes %>%
  left_join(codes, by = "team") %>%
  group_by(year, yearly_game_id) %>%
  mutate(opposing_score = rev(score)) %>%
  mutate(won_by = score - opposing_score) %>%
  ungroup()

```


```{r}
outcomes %>%
  ggplot(aes(score)) +
  geom_histogram(bins = 20) +
  facet_wrap(~ win_status)

  
outcomes %>%
  filter(year == 2019) %>%
  count(round, sort = T)


outcomes %>%
  filter(year == 2019, 
         team %in% c("USA", "NED"))

# of the 3 games each country plays in the "group" round, hwo much did they win by on average

avg_group_scores <- outcomes %>%
  filter(round == "Group") %>%
  group_by(year, team) %>%
  summarize(avg_group_score = mean(score),
            avg_group_won_by = mean(won_by)) %>%
  ungroup()

outcomes %>%
  inner_join(avg_group_scores, by = c("year", "team")) %>%
  filter(round == "Final") %>%
  ggplot(aes(country, avg_group_won_by, fill = win_status)) +
  geom_col() +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_wrap(~ year, scales = "free_x") +
  labs(fill = "Win Status", x = "", y = "Average group win by",
       title = "Does group average scores predict the winner of finals?",
       subtitle = "Yes in all years except 2011. (2015 bad been tied on average score)") +
  guides(fill = guide_legend(reverse = T), element_text(face = "bold", size = 16)) +
  theme(legend.position = "bottom",  
        legend.text = element_text(face = "italic"),
        legend.direction = "horizontal",
        text = element_text(family = "Space Mono"),
        panel.grid = element_blank(),
        plot.title = element_text(size = 20, face = "bold"),
        plot.subtitle = element_text(size = 10, face = "italic"))


```

Applied a simple formula: take the average $ of goals each team had won by in the group rounds. Then take the difference in those numbers, and predict that that's the difference in the playoff scores. 

```{r}
outcomes %>%
  inner_join(avg_group_scores, by = c("year", "team")) %>%
  filter(round != "Group") %>%
  group_by(year, yearly_game_id) %>%
  mutate(difference = diff(avg_group_won_by)) %>%
  ungroup() %>%
  filter(team_num == 2) %>%
  mutate(round = fct_reorder(round, round, length)) %>%
  ggplot(aes(difference, won_by)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(color = "red", linetype=2) +
  facet_wrap(~ round, scales = "free_x") +
  labs(x = "Difference in the team's average group score from their opponent",
       y = "How much they won by",
       title = "Use team performance to predict final performance",
       subtitle = "Use each team performacne in group round to predict their performance in semi-final, quarter final, round of 16, third palce and final")


```

### play by play data

pull football event from StatsBombFreeEvents, Data source: [StatsBomb](https://statsbomb.com)

```{r}
all_matches <- FreeMatches(FreeCompetitions())

wwc_matches <- all_matches %>%
  filter(competition.competition_name == "Women's World Cup")

extract_matches <- function(index) {
  tbl_df(get.matchFree(wwc_matches[index, ])) %>%
    mutate(timestamp = as.difftime(timestamp))
}

matches <- map_df(52, extract_matches, .id = "match_index")

plays <- matches %>%
  inner_join(wwc_matches, by = "match_id") %>%
  transmute(match_id, index, location, possession, minute, second, duration, type = fct_lump(type.name, n = 6),
            player = player.name, position = position.name) %>%
  mutate(x = map_dbl(location, 1, .default = NA),
         y = map_dbl(location, 2, .default = NA),
         timestamp = minute * 60 + second) 

plays %>%
  mutate(player = glue::glue("{ player } ({ position })")) %>%
  filter(!is.na(player)) %>%
  mutate(player = fct_lump(player, 9)) %>%
  filter(player != "Other") %>%
  count(player, type, position, sort = T) %>%
  mutate(type = reorder_within(type, n, player)) %>%
  ggplot(aes(type, n)) +
  geom_col(aes(fill = position)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  coord_flip() +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  facet_wrap(~ player, scales = "free") +
  labs(x = "", y = "",
       title = "Which players were involved in the most plays in 2019 Women's Final",
       fill = "Position") +
  theme(text = element_text(family = "Space Mono"),
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        legend.position = "bottom")

```








### Appendix:

1. how to scrap country code use rvest

```{r}
country_code <- read_html("https://simple.wikipedia.org/wiki/List_of_FIFA_country_codes") %>%
  html_nodes("table") %>%
  map(html_table, fill = TRUE) %>%
  .[2:5] %>%
  bind_rows() %>%
  tbl_df() %>%
  select(Country = Country, team = Code)


```


2. how to reordering categories within ggplot2 facets

[Ordering Categories within ggplot2 Facets](https://trinkerrstuff.wordpress.com/2016/12/23/ordering-categories-within-ggplot2-facets/)


### Summary

Hours: *4 hours*
Resources: 
1. *[HOW TO CREATE PLOTS WITH BEAUTIFUL ANIMATION IN R](https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/)*
2. *[Tidy Tuesday screencast: analyzing Women's World Cup data](https://www.youtube.com/watch?v=ZOQSuapvHqA)*
3. *[REORDERING AND FACETTING FOR GGPLOT2](https://juliasilge.com/blog/reorder-within/)*











