---
title: "ff_data_vis"
author: "Yifei Liu"
date: "5/31/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}

library(tidyverse)
library(tidyquant)
library(ggrepel)
library(ggtext)
library(readxl)
library(transformr) # devtools::install_github("thomasp85/transformr")
detach("package:dplyr", unload = TRUE)
library(dplyr)
theme_set(theme_minimal())

```

```{r}
ff_vol = read_csv("/Users/yifeiliu/Documents/R/data/EDA/datavis/ff_trade_vol.csv") %>%
  janitor::clean_names() %>%
  select(date = date, ff = volume_us_billions)%>%
  mutate(date = mdy(date))

on_bf_vol <- tq_get(x = c("OBFRVOL","EFFRVOL"), 
                    get = "economic.data") %>%
  pivot_wider(names_from = symbol, values_from = price) %>%
  rename(on_bf = OBFRVOL, ff = EFFRVOL)

fund_vol = full_join(on_bf_vol, ff_vol) %>%
  pivot_longer(-date, names_to = "name", values_to = "vol")


fund_vol %>%
  mutate(time = as.numeric(date) / 365.25 + 1970) %>%
  ggplot(aes(time, vol, group = name, color = name)) +
  geom_point(alpha = 0.06) +
  geom_smooth() +
  expand_limits(y = 0) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "<b>Overnight Bank Funding and Effective / Federal Funds Volume</b><br>",
        subtitle = "Combine FF and EFF data together, since FRED don't provide FF data.<p></span><br>*Unit: Billion of U.S. Dollar*<br></p><p><span style = 'color:red;'>Federal Fund Volume </span> <span style = 'color:blue;'>Overnight Bank Funding Volume</span></p>",
       y = "", x = "", legend = "",
       caption = "Data Source: FRED, New York Fed | Yife Liu") +
  theme(plot.title.position = "plot",
    plot.title = element_textbox_simple(
      size = 13,
      lineheight = 1
    ),
    plot.subtitle = element_textbox_simple(
      size = 10,
      lineheight = 1,
      padding = margin(5.5, 5.5, 5.5, 5.5),
    ),
    legend.position = "none",
    plot.caption = element_text(face ="italic"),
    plot.caption.position = "plot")  + 
 scale_color_manual(labels = c("Federal Fund", "Overnight Bank Funding"),
                     values = c("ff" = "red", "on_bf" = "blue"))



```



ON RRP and IOER rate

```{r}
file_name <- dir(path = "/Users/yifeiliu/Documents/R/data/EDA/datavis/RRP", pattern = "rp_rrp_data")

files_path <- paste0("/Users/yifeiliu/Documents/R/data/EDA/datavis/RRP/", file_name)

x = map(files_path, ~read_xls(path =., range = cell_rows(4:5)))

RRP <- do.call(rbind.data.frame, x) %>%
  janitor::clean_names() %>%
  select(date = trade_date, RRP = repo_rate_in_percent) %>%
  mutate(date = ymd(date))

IOER <- tq_get(x = "IOER", get ="economic.data", from = "2010-08-01") %>%
  mutate(date = ymd(date)) %>%
  select(date, IOER = price)

fed_corridor <- left_join(RRP, IOER, by = "date") %>%
  pivot_longer(-date, names_to = "type", values_to = "rate")

IOER_rest <- IOER %>%
  filter(date > ymd("2018-01-02")) %>%
  mutate(type = "IOER") %>%
  rename(rate = IOER)

fed_corridor <- rbind(fed_corridor, IOER_rest)

fed_corridor %>%
  ggplot(aes(date, rate, color = type)) +
  geom_line() +
  labs(title = "<b>ON RRP and IOER Rate</b><br>", x= "", y= "",
       subtitle = "ON RRP rate released two years after each quarterly transaction period<br></p><p></span><br>*Unit: Percentage*<br></p><p><span style = 'color:red;'>ON RRP </span> <span style = 'color:blue;'>IOER</span></p>", caption = "Data Source: New York Fed, FRED | Author: Yifei Liu") +
  scale_color_manual(labels = c("Federal Fund", "Overnight Bank Funding"),
                     values = c("IOER" = "blue", "RRP" = "red")) +
  theme(plot.title.position = "plot",
        legend.position = "none",
        plot.title = element_textbox_simple(
      size = 13,
      lineheight = 1,
      padding = margin(4, 4, 4, 4)
    ),
    plot.subtitle = element_textbox_simple(
      size = 8,
      lineheight = 0,
      padding = margin(5.5, 5.5, 5.5, 5.5)
    ))

```


on GCF repo rate volatility vis

```{r}

gcf_rate <- read_xlsx(path = "/Users/yifeiliu/Documents/R/data/EDA/datavis/RRP/gcf_index_graph.xlsx", sheet = 'Data 2005-2019', range = cell_rows(7:3756)) %>%
  setNames(c("date","MBS", "treasury", "average")) %>%
  mutate(date = ymd(date), type = "GCF") %>%
  select(date, rate = treasury, type)


rbind(gcf_rate, fed_corridor) %>%
  filter(date > ymd("2010-08-04")) %>%
  ggplot(aes(date, rate, color = type)) +
  geom_line() +
  scale_color_manual(labels = c("Federal Fund", "Overnight Bank Funding", "GCF"),
                     values = c("IOER" = "blue", "RRP" = "red", "GCF" = "purple")) +
  labs(title = "<b>ON RRP, IOER Rate and GCF Repo</b><br>", x= "", y= "",
       subtitle = "Treasury GCF Repo® Weighted Average Rate<br></p><p></span><br>*Unit: Percentage*<br></p><p><span style = 'color:red;'>ON RRP</span> <span style = 'color:blue;'>IOER</span></span> <span style = 'color:purple;'>GCF</span></p>", caption = "Data Source: New York Fed, FRED, GTCC | Author: Yifei Liu") +
  theme(plot.title.position = "plot",
        legend.position = "none",
        plot.title = element_textbox_simple(
      size = 13,
      lineheight = 1,
      padding = margin(4, 4, 4, 4)
    ),
    plot.subtitle = element_textbox_simple(
      size = 8,
      lineheight = 0,
      padding = margin(5.5, 5.5, 5.5, 5.5)
    ))


```


## LIBOR Rate

How Basel III break Libor rate US dollar

```{r}

libor_rate <- c("USDONTD156N", "USD1WKD156N", "USD1MTD156N", "USD2MTD156N", "USD3MTD156N", "USD6MTD156N", "USD12MD156N")

libor_frame <- tibble(name = c("USDONTD156N", "USD1WKD156N", "USD1MTD156N", "USD2MTD156N", "USD3MTD156N", "USD6MTD156N", "USD12MD156N"),
                          list = c("on", "1w", "1m","2m", "3m", "6m", "12m"))

libor_series <- tq_get(x = libor_rate, get = "economic.data", from = "2011-01-01")

x = libor_series %>%
  full_join(libor_frame, by = c("symbol" = "name")) %>%
  mutate(date = ymd(date),
         list = factor(list, levels = c("on", "1w", "1m","2m", "3m", "6m", "12m"))) %>%
  filter(wday(date) == 5) %>%
  select(date, price, name = list)


x %>%
  ggplot(aes(date, price, group = date, color = date)) +
  geom_line() +
  scale_fill_gradientn(colours = terrain.colors(10)) +
  transition_time(date) +
  ease_aes('linear') +
  labs(title = 'Date: {frame_time}')



```

 
 
LIBOR-ISO gap

this rate measure the interbank lending stability. Some bank have access to Fed Fund rate, and some don't. And When some of these bank couldn't find counterpart who are willing to lend them at fed fund rate, libor rate increase.  This rate raise to highest when Bear Stearn failed. 

```{r}

libor_ois <- tq_get(x = c("FEDFUNDS","USD1MTD156N"), 
                    get = "economic.data", from = "2007-07-01",
                    to = "2011-01-01") %>%
  pivot_wider(names_from = symbol, values_from = price) %>%
  rename(libor = USD1MTD156N, ff = FEDFUNDS)

libor_ois %>%
  mutate(libor = case_when(is.na(libor) ~ dplyr::lag(libor),
                     TRUE ~ libor)) %>%
  filter(lubridate::day(date) == 1) %>%
  mutate(rate_gap = libor - ff) %>%
  ggplot(aes(date, rate_gap)) +
  geom_line()

libor_ois %>%
  mutate(libor = case_when(is.na(libor) ~ dplyr::lag(libor),
                     TRUE ~ libor)) %>%
  filter(lubridate::day(date) == 1) %>%
  mutate(rate_gap = libor - ff) %>%
  ggplot(aes(date, rate_gap)) +
  geom_line()


```






