---
title: "Purrr walkthrough"
author: "Yifei Liu"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=F, warning=F}
library(tidyverse)
library(purrr)
library(magrittr)
library(rlang)
library(gapminder)
library(repurrrsive)
library(listviewer)
library(listviewer)
theme_set(theme_minimal())

detach("package:dplyr", unload = TRUE)
library(dplyr)

```

## Vectors and lists

### Atomic vectors

```{r}
(v_log <- c(TRUE, TRUE, FALSE, FALSE))
is.logical(v_log)
is_logical(v_log)

```

### Coercion

### List

### List indexing

```{r}
(x <- list(1:3, c("four", "five")))

x[c(F, T)]


```

### Vectorized operations

```{r}
n <- 5
seq_len(n) ^ 2

```


## Relationship to base and plyr functions

### Why not base?

### Why purrr?

#### lapply() vs. purrr::map()

```{r}
map(got_chars[1:3], "name")

```

#### \_(ツ)_/¯ vs. map_dfr()

map_dfr function is list in data frame out

```{r}
map_dfr(got_chars[23:25], `[`, c("name", "playedBy"))

tibble::tibble(name = map_chr(got_chars, "name"),
               id = map_int(got_chars, "id"))


```


#### mapply() vs. map2(), pmap()

when you need to iterate over 2 or more vectors/lists in parrallel, the base options is mapply

pmap_*() applied .f to each row

```{r}
nms <- got_chars[16:18] %>%
  map_chr("name")
birth <- got_chars[16:18] %>%
  map_chr("born")

map2_chr(nms, birth, ~paste(.x, "was born", .y))

df <- tibble::tibble(
  nms,
  connector = "was born",
  birth
)

pmap_chr(df, paste)

```


#### aggregate() vs. dplyr::summarize()

```{r}
(mini_gap <- gapminder %>%
   filter(country %in% c("Canada", "Germany"),
          year > 2000))

mini_gap %>%
  group_by(country) %>%
  summarize(lifeExp = mean(lifeExp))

gapminder %>%
  group_by(country) %>%
  summarize(cor = cor(lifeExp, year))

```


#### by() vs. tidyr::nest()

```{r}
nested_df <- gapminder %>%
  group_by(country, continent) %>%
  nest() %>%
  mutate(fit = map(data, ~lm(lifeExp ~ year, data = .x)))

glimpse(nested_df,max.level = 1)

nested_df$fit[[1]]

nested_df %>%
  filter(continent == "Oceania") %>%
  .$fit

nested_df %>%
  mutate(coefs = map(fit, coef),
         intercept = map_dbl(coefs, 1),
         slope = map_dbl(coefs, 2)) %>%
  select(- coefs)

```

## Explore the example lists: Wes Anderson, Game of Thrones, GitHub

### Inspect and explore

### Wes Anderson color palettes

```{r}
str(wesanderson)

```

### Game of Thrones POV characters

```{r}
str(got_chars, list.len = 3)
dplyr::glimpse(got_chars, list.len = 3)

```

## Introduction to map(): extract elements

### Vectorized and “list-ized” operations

### Type-specific map


```{r}

x <- map(got_chars, `[`, c("name", "culture", "gender", "born"))

str(x[16:18])

x <- map(got_chars, extract, c("name", "culture", "gender", "born"))
str(x[16:20], max.level = 2, list.len = 2)

map_chr(got_chars[[1]], class)

```

### Data frame output

```{r}
map_dfr(got_chars, extract, c("name", "culture", "gender", "id", "born", "alive"))

```

## Simplifying data from a list of GitHub users

### Get several GitHub users

```{r}
str(gh_users, max.level = 2, list.len = 3)


```


### Name and position shortcuts

### Vector input to extraction shortcuts

```{r}
gh_repos %>%
  map_chr(c(1, 4, 1))

gh_repos[[1]][[1]]$owner[[1]]

str(gh_repos, list.len = 3)

```
IMPORTANT: Note that this does NOT give elements 1 and 3 of gh_repos. It extracts the first repo for each user and, within that, the 3rd piece of information for the repo.


### List inside a data frame


```{r}
(unames <- map_chr(gh_repos, c(1,4,1)))



(udf <- gh_repos %>%
    set_names(unames) %>%
    enframe("username", "gh_repos"))


udf %>%
  mutate(n_repos = map_int(gh_repos, length))

one_user <- udf$gh_repos[[1]]

identical(one_user, gh_repos[[1]])

str(one_user, max.level = 3, list.len = 5)

one_repo <- one_user[[1]]

identical(one_repo, gh_repos[[1]][[1]])

str(one_repo)

map_df(one_user, extract, c("name", "fork", "open_issues"))

udf %>%
  mutate(repo_info = gh_repos %>%
           map(. %>% map_df(extract, c("name", "fork", "open_issues"))))

rdf <- udf %>%
  mutate(repo_info = gh_repos %>%
           map(. %>% map_df(extract, c("name", "fork", "open_issues")))) %>%
  select(-gh_repos) %>%
  unnest()

rdf %>%
  filter(!fork) %>%
  select(-fork) %>%
  group_by(username) %>%
  arrange(username, desc(open_issues)) %>%
  slice(1:3)

```

## Specifying the function in map() + parallel mapping

### map() function specification

```{r}
aliases <- set_names(map(got_chars, "aliases"), map_chr(got_chars, "name"))

aliases

```


### Existing functio

```{r}
my_function <- function(x) paste(x, collapse = " | ")

map(aliases, my_function)

```

### Anonymous function, conventional

```{r}
map(aliases, function(x) paste(x, collapse = " | "))

map(aliases, paste, collapse = " | ")

```


### Anonymous function, formula

```{r}
map(aliases, ~ paste(.x, collapse = " | "))

```


### Workflow advice

```{r}
(a <- map(got_chars, "aliases")[[19]])
(a <- map(got_chars, "aliases")[[16]])

paste(a, sep = " | ")
paste(a, collapse  = " | ")

got_chars[15:17] %>%
  map("aliases") %>%
  map_chr(paste, collapse = " | ")


```

### List to data frame

```{r}
aliases <- set_names(map(got_chars, "aliases"), map_chr(got_chars, "name"))

map_chr(aliases, ~paste(.x, collapse = " | ")) %>%
  enframe(value = "aliases")



```

### Exercises

1. Create a list allegiances that holds the characters’ house affiliations.

```{r}
allegiances <- set_names(map(got_chars, "allegiances"), map_chr(got_chars, "name"))

```

2. Create a character vector nms that holds the characters’ names.

```{r}
nms <- map_chr(got_chars, "name")

```

3. Apply the names in nms to the allegiances list via set_names.

```{r}
allegiances <- set_names(map(got_chars, "allegiances"), map_chr(got_chars, "name"))
```

4. Re-use the code from above to collapse each character’s vector of allegiances down to a string.

```{r}
map(allegiances, ~paste(.x, collapse = " | "))

```

5. We said that any elements passed via ... would be used “as is”. Specifically they are not used in a vectorized fashion. What happens if you pass collapse = c(" | ", " * ")? Why is that?

```{r}
map(allegiances, ~paste(.x, collapse = c(" | ", " * ")))

identical(map(allegiances, ~paste(.x, collapse = " | ")), map(allegiances, ~paste(.x, collapse = c(" | ", " * "))))

map_chr(allegiances, ~ paste(.x, collapse = " | ")) %>%
  enframe(value = "allegiances")

```

the result just ingore the second collapse. 

### Parallel map

#### map2()

```{r}
nms <- map_chr(got_chars, "name")
birth <- got_chars %>%
  map_chr("born")

map2_chr(nms, birth, ~paste(.x, "was born", .y))

```

#### pmap()

```{r}
df <- got_chars %>% {
  tibble::tibble(
    name = map_chr(., "name"),
    aliases = map(., "aliases"),
    allegiance = map(., "allegiances")
  )
}

my_fun <- function(name, aliases, allegiances) {
  paste(name, "has", length(aliases), "aliases and",
        length(allegiances), "allegiances")
}

df %>%
  pmap_chr(my_fun) %>%
  tail()

```

## Trump Android words

```{r}
load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
```

### Trump Android words

```{r}
regex <- "badly|crazy|weak|spent|strong|dumb|joke|guns|funny|dead"

tweets <- trump_tweets_df$text

tweets %>% head() %>% strtrim(70)

tweets[c(1, 2, 5, 6, 198, 347, 919)] %>%
  strtrim(70)

```

### Are you ready for gregexpr()?

```{r}
matches <- gregexpr(regex, tweets)

matches[[7]]

trump_tweets_df %>%
  filter(str_detect(text, regex))


map(matches, ~attr(.x, which = "match.length")) %>%
  head()

lengths(matches)

map(matches, function(x) sum(x > 0)) %>%
  head()

map_int(matches, ~sum(.x > 0)) 

match_length <- map(matches, attr, which = "match.length")

tibble(
  naive_length = lengths(matches),
  n_words = map_int(matches, ~sum(.x > 0))
)

```

### Strip the attributes from matches

```{r}
match_first <- map(matches, as.vector)

head(match_first)

```


### Assess progress in a small example


```{r}
head(match_length)

trump_tweets_df %>%
  mutate(contain = map(tweets, str_extract_all, pattern = regex),
         words = map_chr(contain, ~paste(.x, collapse = " | "))) %>%
  select(- contain) %>%
  mutate(words = case_when(words == "character(0)" ~ NA_character_,
                           T ~ words))

```

## Sample from groups, n varies by group

```{r}
nested_iris <- iris %>%
  group_by(Species) %>%
  nest() %>%
  mutate(n = c(2,5,3))

nested_iris

sampled_iris <- nested_iris %>%
  mutate(samp = map2(data, n, sample_n))

sampled_iris

sampled_iris %>%
  select(Species, samp) %>%
  unnest()

iris %>%
  group_by(Species) %>%
  nest() %>%
  mutate(n = c(2,5,3),
         samp = map2(data, n, sample_n)) %>%
  select(Species, samp) %>%
  unnest()

```





Reference: [purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html)






























